---
title: "Hankowsky_HW7"
output: html_document
date: "2022-10-19"
---

```{r setup, include=FALSE}
library(Sleuth3)
library(tidyverse)
library(asht)
library(mosaic)
library(GGally)
library(gridExtra)
library(olsrr)
library(car)

```

Hankowsky Homework Solutions 


# Basic Problems 
## 10.31

```{r}
#read in the data 
parental_care <- Sleuth3::ex1031


#looking at the dataframe
head(parental_care)


#plot the clutch volume by mass 
parental_care %>%
  ggplot() + 
  geom_point(aes( x = BodyMass, y =  ClutchVolume, shape = Group, color = Group)) + 
  scale_color_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  scale_shape_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  labs(title = "Clutch Volume v Body Mass", 
       x = "Body Mass (kg)", y = bquote("Clutch Volume " (mm^3))) + 
  theme_classic()


#transforming the data as described in the question 
parental_care <- parental_care %>%
  mutate(log_clutch = log(ClutchVolume), 
         log_mass = log(BodyMass), 
         Group = relevel(Group, ref = "Mani"))


#plot the clutch volume by mass 
parental_care %>%
  ggplot() + 
  geom_point(aes(x = log_mass, y =  log_clutch, shape = Group, color = Group)) + 
  scale_color_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  scale_shape_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  labs(title = "Log(Clutch Volume) v Log(Body Mass)", 
       x = "Log(Body Mass (kg))", y = bquote("Log(Clutch Volume "~(mm^3)~")")) + 
  theme_classic()


#run and compare the regression models 
parental_care_lm1 <- lm(log_clutch ~ log_mass + Group + log_mass:Group, data = parental_care)
parental_care_lm2 <- lm(log_clutch ~ log_mass + Group, data = parental_care)
parental_care_lm3 <- lm(log_clutch ~ log_mass, data = parental_care)

anova(parental_care_lm2, parental_care_lm1) #testing for interaction effects 
anova(parental_care_lm3, parental_care_lm2) #testing for Group effects


#run the optimal regression
parental_care_lm <- lm(log_clutch ~ log_mass + Group + log_mass:Group, data = parental_care)
summary(parental_care_lm)


#back-transform the estimates 
exp(parental_care_lm$coefficients)
exp(confint(parental_care_lm))

100*(exp(parental_care_lm$coef) - 1) 
100*(exp(confint(parental_care_lm)) - 1)


#plot the diagnostics 
par(mfrow=c(2,2))
plot(parental_care_lm)


#plot the presentation graphic 
par(mfrow=c(1,1))
parental_care %>%
  ggplot(aes(x = log_mass, y =  log_clutch, shape = Group, color = Group)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = F) +
  scale_color_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  scale_shape_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  labs(title = "Log(Clutch Volume) v Log(Body Mass)", 
       x = "Log(Body Mass (kg))", y = bquote("Log(Clutch Volume "~(mm^3)~")")) + 
  theme_classic()

```

A scatterplot of the raw clutch volume as a function of body mass data showed the need for a transformation. A log-transformation was preformed and the scatterplot showed no problems, so the analysis was conducted with the log-transformed data. There is strong evidence that the relationship between clutch volume and body mass differed between the Mani group and the other groups (Extra Sum of Squares F-Test, p-value < 0.001). 




***

## 11.25

```{r}
#read in the data 
deposit_feeders <- Sleuth3::ex1125

#looking at the dataframe
head(deposit_feeders)

#transforming the data
deposit_feeders <- deposit_feeders %>%
  mutate(log_weight = log(Weight +1), 
         log_ingestion = log(Ingestion +1), 
         log_organic = log(Organic +1))

#looking at scatterplots of the variables
ggpairs(deposit_feeders, columns = 6:8)

#run the mutiple regression 
deposit_feeders_lm1 <- lm(log_ingestion ~ log_weight + log_organic + Bivalve, data = deposit_feeders)
summary(deposit_feeders_lm1)


#plot the diagnostics 
par(mfrow=c(2,2))
plot(deposit_feeders_lm1)

plot(deposit_feeders_lm1, which = 4)


#plot the dfbetas plot
#plotting it dr gallaghers way
deposit_feeders_lm1_dfbs <- dfbetas(deposit_feeders_lm1)
head(deposit_feeders_lm1_dfbs)

plot(deposit_feeders_lm1_dfbs[ , c("log_weight", "log_organic")], 
     main="dfbetas plot")
identify(deposit_feeders_lm1_dfbs[ ,"log_weight"],deposit_feeders_lm1_dfbs[ ,"log_organic"])


#plotting it using the olsrr::ols_plot_dfbetas
ols_plot_dfbetas(deposit_feeders_lm1)


#re-run the regression without observation 22
#creating an ID number for each row
deposit_feeders <- deposit_feeders %>%
  mutate(ID = 1:nrow(deposit_feeders)) %>%
  relocate(ID)

#re-run regression 
deposit_feeders_lm2 <- update(deposit_feeders_lm1, ~ ., subset = (ID != 20 & ID != 22))

#plot the diagnostics of the updated regression 
par(mfrow=c(2,2))
plot(deposit_feeders_lm2)

plot(deposit_feeders_lm2, which = 4)

#plot the dfbetas plot
ols_plot_dfbetas(deposit_feeders_lm2)



#looking at teh bivalve effect
deposit_feeders_lm3 <- lm(log_ingestion ~ log_weight + log_organic, data = deposit_feeders, 
                          subset = (ID != 20 & ID != 22))

anova(deposit_feeders_lm3, deposit_feeders_lm2) #testing for bivalve effects 


#looking at the final model 
summary(deposit_feeders_lm3)

#compoenent residual plots 
crPlots(deposit_feeders_lm2)

#back transforming the data 
exp(deposit_feeders_lm2$coefficients)
exp(confint(deposit_feeders_lm2))

100*(exp(deposit_feeders_lm2$coefficients) - 1) 
100*(exp(confint(deposit_feeders_lm2)) - 1)


```
There is strong evidence that the relationship between ingestion rate and the percentage of organic matter in food and animal weight differed between bivalves and non-bivalves (Extra Sum of Squares F-Test, p-value < 0.001). However, two of the bivalve observations were considerable outliers and were removed from the analysis. Thus, the bivalve term is only being fitted to one observation and I did not think it was appropriate to fit a model with a bivalve term if there was only one bivalve observation in the dataset.

These data provide overwhelming evidence that the ingestion rate is associated with the percentage of organic matter in food, after accounting for animal weight (p-value < 0.0001, multiple regression). 



***

# Supplemental Problems 

## 10.32

```{r}
#read in the data 
height <- Sleuth3::ex0726


#looking at the dataframe
head(height)


#transposing the data 
height_long <- height %>%
  pivot_longer(cols = c("Father", "Mother"), 
               names_to = "parent", 
               values_to = "parent_height")
  

#plot
height_long %>%
  ggplot() + 
  geom_point(aes(x = parent_height, y =  Height)) + 
  facet_grid(Gender ~ parent) + 
  labs(x = "Parent's Height (in)", y = "Child's Adult Height (in)")


#regression 
height_lm <- lm(Height ~ Mother + Father + Gender, data = height)
summary(height_lm)


#plot the diagnostics 
par(mfrow=c(2,2))
plot(height_lm)



```


The equation for predicting a  child’s adult height from their mother’s height, their father’s height, and their gender is y = 16.43 + 0.32(Mother's Height) + 0.39(Father's Height) + 5.22(Male Gender). 


***

## 12.20

```{r}
#read in the data 
galapagos <- Sleuth3::ex1220

#looking at the dataframe
head(galapagos)

#looking at scatterplots of the variables
ggpairs(galapagos, columns = 2:8)


#fit a model with all the variables 
galapagos_lm1 <- lm(Total ~ Area, data = galapagos)
summary(galapagos_lm1)

galapagos_lm2 <- lm(Total ~ Area + Elev, data = galapagos)
summary(galapagos_lm2)

galapagos_lm3 <- lm(Total ~ Area + DistSc, data = galapagos)
summary(galapagos_lm3)

galapagos_lm4 <- lm(Total ~ Area + AreaNear, data = galapagos)
summary(galapagos_lm4)


galapagos_lm2 <- lm(Total ~ Area + Elev + DistSc + AreaNear, data = galapagos)
summary(galapagos_lm2)


```




***

# Master Problem 

