---
title: "Hankowsky_HW7"
output: html_document
date: "2022-10-19"
---

```{r setup, include=FALSE}
library(Sleuth3)
library(tidyverse)
library(asht)
library(mosaic)
library(GGally)
library(gridExtra)
library(olsrr)
library(car)

```

Hankowsky Homework Solutions 


# Basic Problems 
## 10.31
Who Looks After the Kids? Different bird species have different strategies: Maternal, Paternal,
and BiParental care. In 1984 J. Van Rhijn argued in the Netherlands Journal of Zoology
that parental care was the ancestral condition, going back to the dinosaur predecessors of birds.
D. J. Varricchio et al. (Science, Vol. 322, Dec. 19, 2008, pp. 1826–28) tested that argument by comparing
the relationships between Clutch Volume and adult Body Mass in six different groups: modern
maternal-care bird species (MatI n D 171), modern paternal-care bird species (PatI n D 40), modern
biparental-care bird species (BiPI n D 204), modern maternal-care crocodiles (CrocI n D 19),
non-avian maniraptoran dinosaurs thought to be ancestors of modern birds (ManiI n D 3), and other
non-avian dinosaurs (OthrI n D 6). The question of interest was which group of modern creatures
most closely matches the relationship in the maniraptoran dinosaurs. A partial listing of the data
appears in Display 10.27.
Fit a single model, with clutch volume (possibly transformed) as the response, that allows for
separate straight lines in each group, using indicator variables for the different groups. For example,
by selecting Mani as the reference group, you can determine how close any other group is to Mani
by dropping out the indicator for that group and its product with the body mass variable. Do this for
all other groups. (a) Are there differences among the relationships in the groups other than the Mani
group? What if there were none? (b) What do you conclude about which other group is nearest the
maniraptoran dinosaur group? (c) Is the model defensible? (d) Comment on the study design.
```{r}
#read in the data 
parental_care <- Sleuth3::ex1031


#looking at the dataframe
head(parental_care)


#plot the clutch volume by mass 
parental_care %>%
  ggplot() + 
  geom_point(aes( x = BodyMass, y =  ClutchVolume, shape = Group, color = Group)) + 
  scale_color_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  scale_shape_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  labs(title = "Clutch Volume v Body Mass", 
       x = "Body Mass (kg)", y = bquote("Clutch Volume " (mm^3))) + 
  theme_classic()


#transforming the data as described in the question 
parental_care <- parental_care %>%
  mutate(log_clutch = log(ClutchVolume), 
         log_mass = log(BodyMass), 
         Group = relevel(Group, ref = "Mani"))


#plot the clutch volume by mass 
parental_care %>%
  ggplot() + 
  geom_point(aes(x = log_mass, y =  log_clutch, shape = Group, color = Group)) + 
  scale_color_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  scale_shape_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  labs(title = "Log(Clutch Volume) v Log(Body Mass)", 
       x = "Log(Body Mass (kg))", y = bquote("Log(Clutch Volume "~(mm^3)~")")) + 
  theme_classic()


#run and compare the regression models 
parental_care_lm1 <- lm(log_clutch ~ log_mass + Group + log_mass:Group, data = parental_care)
parental_care_lm2 <- lm(log_clutch ~ log_mass + Group, data = parental_care)
parental_care_lm3 <- lm(log_clutch ~ log_mass, data = parental_care)

anova(parental_care_lm2, parental_care_lm1) #testing for interaction effects 
anova(parental_care_lm3, parental_care_lm2) #testing for Group effects


#run the optimal regression
parental_care_lm <- lm(log_clutch ~ log_mass + Group, data = parental_care)
summary(parental_care_lm)


#back-transform the estimates 
exp(parental_care_lm$coefficients)
exp(confint(parental_care_lm))

100*(exp(parental_care_lm$coef) - 1) 
100*(exp(confint(parental_care_lm)) - 1)


#plot the diagnostics 
par(mfrow=c(2,2))
plot(parental_care_lm)


#plot the presentation graphic 
par(mfrow=c(1,1))
parental_care %>%
  ggplot(aes(x = log_mass, y =  log_clutch, shape = Group, color = Group)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = F) +
  scale_color_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  scale_shape_discrete(breaks=c("BiP", "Croc", "Mani", "Mat", "Othr", "Pat"),
                      labels=c("Modern Biparental-Care \n Bird Species \n", 
                                "Modern Maternal-Care \n Crocodiles \n", 
                                "Non-Avain Maniraptoran \n Dinosaurs \n", 
                                "Modern Maternal-Care \n Bird Species \n", 
                                "Other Non-Avain \n Dinosaurs \n", 
                                "Modern Paternal-Care \n Bird Species \n")) + 
  labs(title = "Log(Clutch Volume) v Log(Body Mass)", 
       x = "Log(Body Mass (kg))", y = bquote("Log(Clutch Volume "~(mm^3)~")")) + 
  theme_classic()

```

Adjusted for body mass, the median clutch volume for Non-Avain Maniraptoran Dinosuars is 65% greater than the median clutch volumne for Modern Biparental-Care Bird Species (95% confidence interval: 34% to 82%). 


***

## 11.25
Ingestion Rates of Deposit Feeders. The ingestion rates and organic consumption percentages
of deposit feeders were considered in Exercise 9.21. The data set ex1125 repeats these data, but
this time with three additional bivalve species included (the last three). The researcher wished to see
if ingestion rate is associated with the percentage of organic matter in food, after accounting for animal
weight, but was unsure about whether bivalves should be included in the analysis. Analyze the
data to address this question of interest. (Data from L. M. Cammen, “Ingestion Rate: An Empirical
Model for Aquatic Deposit Feeders and Detritivores,” Oecologia 44 (1980): 303–10.)
```{r}
#read in the data 
deposit_feeders <- Sleuth3::ex1125

#looking at the dataframe
head(deposit_feeders)

#transforming the data
deposit_feeders <- deposit_feeders %>%
  mutate(log_weight = log(Weight +1), 
         log_ingestion = log(Ingestion +1), 
         log_organic = log(Organic +1))

#looking at scatterplots of the variables
ggpairs(deposit_feeders, columns = 6:8)

#run the mutiple regression 
deposit_feeders_lm1 <- lm(log_ingestion ~ log_weight + log_organic + Bivalve, data = deposit_feeders)
summary(deposit_feeders_lm1)


#plot the diagnostics 
par(mfrow=c(2,2))
plot(deposit_feeders_lm1)

plot(deposit_feeders_lm1, which = 4)


#plot the dfbetas plot
#plotting it dr gallaghers way
deposit_feeders_lm1_dfbs <- dfbetas(deposit_feeders_lm1)
head(deposit_feeders_lm1_dfbs)

plot(deposit_feeders_lm1_dfbs[ , c("log_weight", "log_organic")], 
     main="dfbetas plot")
identify(deposit_feeders_lm1_dfbs[ ,"log_weight"],deposit_feeders_lm1_dfbs[ ,"log_organic"])


#plotting it using the olsrr::ols_plot_dfbetas
ols_plot_dfbetas(deposit_feeders_lm1)


#re-run the regression without observation 22
#creating an ID number for each row
deposit_feeders <- deposit_feeders %>%
  mutate(ID = 1:nrow(deposit_feeders)) %>%
  relocate(ID)

#re-run regression 
deposit_feeders_lm2 <- update(deposit_feeders_lm1, ~ ., subset = (ID != 20 & ID != 22))

#plot the diagnostics of the updated regression 
par(mfrow=c(2,2))
plot(deposit_feeders_lm2)

plot(deposit_feeders_lm2, which = 4)

#plot the dfbetas plot
ols_plot_dfbetas(deposit_feeders_lm2)




#looking at the final model 
summary(deposit_feeders_lm2)

#compoenent residual plots 
crPlots(deposit_feeders_lm2)

#back transforming the data 
exp(deposit_feeders_lm2$coefficients)
exp(confint(deposit_feeders_lm2))

100*(exp(deposit_feeders_lm2$coefficients) - 1) 
100*(exp(confint(deposit_feeders_lm2)) - 1)

```




***

# Supplemental Problems 

## 10.32
Galton’s Height Data. Reconsider the data in Exercise 7.26 on heights of adult children and
their parents. Ignore the possible dependence of observations on children from the same family for
now to answer the following: (a) What is an equation for predicting a child’s adult height from their
mother’s height, their father’s height, and their gender? (b) By how much does the male mean height
exceed the female mean height for children whose parents’ heights are the same? (c) Find a 95%
prediction interval for a female whose father’s height is 72 inches and whose mother’s height is 64
inches.
```{r}
#read in the data 
height <- Sleuth3::ex0726


#looking at the dataframe
head(height)


#transposing the data 
height_long <- height %>%
  pivot_longer(cols = c("Father", "Mother"), 
               names_to = "parent", 
               values_to = "parent_height")
  

#plot
height_long %>%
  ggplot() + 
  geom_point(aes(x = parent_height, y =  Height)) + 
  facet_grid(Gender ~ parent) + 
  labs(x = "Parent's Height (in)", y = "Child's Adult Height (in)")


#regression 
height_lm <- lm(Height ~ Mother + Father + Gender, data = height)
summary(height_lm)


#plot the diagnostics 
par(mfrow=c(2,2))
plot(height_lm)



```


The equation for predicting a  child’s adult height from their mother’s height, their father’s height, and their gender is y = 16.43 + 0.32(Mother's Height) + 0.39(Father's Height) + 5.22(Male Gender). 


***

## 12.20
Galapagos Islands. The data in Display 12.17 come from a 1973 study. (Data from M. P.
Johnson and P. H. Raven, “Species Number and Endemism: The Galapagos Archipelago Revisited,”
Science 179 (1973): 893–5.) The number of species on an island is known to be related to the island’s
area. Of interest is what other variables are also related to the number of species, after island area
is accounted for, and whether the answer differs for native species and nonnative species. (Note:
Elevations for five of the islands were missing and have been replaced by estimates for purposes of
this exercise.)
```{r}
#read in the data 
galapagos <- Sleuth3::ex1220

#looking at the dataframe
head(galapagos)

#looking at scatterplots of the variables
ggpairs(galapagos, columns = 2:8)


#fit a model with all the variables 
galapagos_lm1 <- lm(Total ~ Area + Elev + DistNear + DistSc + AreaNear, data = galapagos)
summary(galapagos_lm1)

galapagos_lm1 <- lm(Total ~ Area + Elev + DistNear + DistSc + AreaNear, data = galapagos)

```




***

# Master Problem 

