---
title: "Hankowsky_HW11"
output: html_document
date: "2022-11-18"
---


```{r setup, include=FALSE}
library(Sleuth3)
library(tidyverse)
library(asht)
library(mosaic)
library(GGally)
library(gridExtra)
library(olsrr)
library(car)
library(Hmisc)
library(MASS)

```

Hankowsky Homework Solutions 


# Basic Problems 
## 22.16

Murder–Suicides by Deliberate Plane Crash. Some sociologists suspect that highly publicized
suicides may trigger additional suicides. In one investigation of this hypothesis, D. P. Phillips
collected information about 17 airplane crashes that were known (because of notes left behind) to
be murder–suicides. For each of these crashes, Phillips reported an index of the news coverage (circulation
in nine newspapers devoting space to the crash multiplied by length of coverage) and the
number of multiple-fatality plane crashes during the week following the publicized crash. The data
are exhibited in Display 22.12. (Data from D. P. Phillips, “Airplane Accident Fatalities Increase Just
After Newspaper Stories AboutMurder and Suicide,” Science 201 (1978): 748–50.) Is there evidence
that the mean number of crashes increases with increasing levels of publicity of a murder–suicide?

```{r}
#read in the data
planes <- Sleuth3::ex2216

#take a look at the data
head(planes)

#plot the data
planes %>%
  ggplot() + 
  geom_point(aes(x = Index, y = Crashes)) + 
  theme_classic()


# REGRESSION ANALYSIS USING BACKWARD ELIMINATION
planes_glm <- glm(Crashes ~ Index, family = poisson, data = planes)
summary(planes_glm)

#back transform the estimates 
exp(cbind(OR = coef(planes_glm), confint(planes_glm)))

#Presentation plot 
planes %>%
  ggplot(aes(x = Index, y = Crashes)) + 
  geom_point() + 
  geom_smooth(method = 'glm', method.args = list(family = 'poisson')) + 
  labs(x = "Index of News Coverage", y = "Number of Plane Crashes") + 
  theme_classic()
  
```

<br>

Associated with each 1 crash increase there is a 0.2% increase in the index of news coverage (95% CI: 0.03 to 0.3%). 

<br>

***



## 22.20

Cancer Deaths of Atomic Bomb Survivors. The data in Display 22.13 are the number of
cancer deaths among survivors of the atomic bombs dropped on Japan during World War II, categorized
by time (years) after the bomb that death occurred and the amount of radiation exposure that
the survivors received from the blast. (Data from D. A. Pierce, personal communication.) Also listed
in each cell is the person-years at risk, in 100’s. This is the sumtotal of all years spent by all persons
in the category. Suppose that the mean number of cancer deaths in each cell is Poisson with mean
! D risk"rate, where risk is the person-years at risk and rate is the rate of cancer deaths per person
per year. It is desired to describe this rate in terms of the amount of radiation, adjusting for the effects
of time after exposure. (a) Using log(risk) as an offset, fit the Poisson log-linear regression model
with time after blast treated as a factor (with seven levels) and with rads and rads-squared treated
as covariates. Look at the deviance statistic and the deviance residuals. Does extra-Poisson variation
seem to be present? Is the rads-squared term necessary? (b) Try the same model as in part (a); but
instead of treating time after bomb as a factor with seven levels, compute the midpoint of each interval
and include log(time) as a numerical explanatory variable. Is the deviance statistic substantially
larger in this model, or does it appear that time can adequately be represented through this single
term? (c) Try fitting a model that includes the interaction of log(time) and exposure. Is the interaction
significant? (d) Based on a good-fitting model, make a statement about the effect of radiation
exposure on the number of cancer deaths per person per year (and include a confidence interval if
you supply an estimate of a parameter).


```{r}
#read in the data
cancer <- Sleuth3::ex2220

#take a look at the data
head(cancer)

#create the variables as dictated in the question 
cancer %>% 
  mutate()


```

<br>


<br>

***



# Supplemental Problems 
## 22.18

Galapagos Islands. Reanalyze the data in Exercise 12.20 with number of native species as the
response, but using log-linear regression. (a) Fit the model with log area, log elevation, log of distance
from nearest island, and log area of nearest island as explanatory variables; and then check for extra-
Poisson variation. (b) Use backward elimination to eliminate insignificant explanatory variables.
(c) Describe the effects of the remaining explanatory variables.

```{r}
#read in the data
islands <- Sleuth3::ex1220

#take a look at the data
head(islands)

#create the variables as dictated in the question 
islands <- islands %>% 
  mutate(log_area = log(Area), 
         log_elevation = log(Elev), 
         log_distnear = log(DistNear), 
         log_areanear = log(AreaNear))

#scatterplot matrix of all the variables of interest 
ggpairs(islands, columns = (9:12))


# PART A
island_glm_a <- glm(Native ~ log_area + log_elevation + log_distnear + log_areanear, family = poisson, data = islands)
summary(island_glm_a)

#deviance goodness-of-fit test
1-pchisq(island_glm_a$deviance, df = island_glm_a$df.residual)

#plot of deviance residuals against the fitted means
gf_point(residuals(island_glm_a, type = "deviance") ~ fitted(island_glm_a)) %>%
  gf_hline(yintercept = 0, color = "blue") %>%
  gf_hline(yintercept = 2, color = "gray60", linetype = 2) %>%
  gf_hline(yintercept = -2, color = "gray60", linetype = 2) %>%
  gf_labs(x = "Fitted Means", y = "Deviance Residuals")



# PART B - using quasipoisson based on Part A 
island_glm_b1 <- glm(Native ~ log_area + log_elevation + log_distnear + log_areanear, family = quasipoisson, data = islands)
summary(island_glm_b1)

island_glm_b2 <- update(island_glm_b1, ~ . - log_elevation)
summary(island_glm_b2)

island_glm_b3 <- update(island_glm_b2, ~ . - log_distnear)
summary(island_glm_b3)


island_glm_optimal <- glm(Native ~ log_area + log_areanear, family = quasipoisson, data = islands)
summary(island_glm_optimal)

#back transform the estimates 
exp(cbind(OR = coef(island_glm_optimal), confint(island_glm_optimal)))

```

<br>

The deviance goodness of fit test provides very strong evidence that there is extra-Poisson variation (p-value < 0.0001). The plot of deviance residuals against the fitted means confirms that it is not a result of one or two outliers, rather it seems that the variability in the response exceeds that which is allowed for by the Poisson log-linear model. Thus, Part B was conducted using quasi-likelihood methods. The backwards regression analysis found no evidence for an effect of log[Elevation] (p-value = 0.67) or log[Distance to the Nearest Island] (p-value = 0.13). 

<br>

***



## 22.24

Valve Failure in Nuclear Reactors. Display 22.16 shows characteristics and numbers of failures
observed in valve types from one pressurized water reactor. There are five explanatory factors:
system (1 D containment, 2 D nuclear, 3 D power conversion, 4 D safety, 5 D process auxiliary);
operator type (1 D air, 2 D solenoid, 3 D motor-driven, 4 D manual); valve type (1 D ball, 2 D
butterfly, 3 D diaphragm, 4 D gate, 5 D globe, 6 D directional control); head size (1 D less than 2
inches, 2 D 2–10 inches, 3 D 10–30 inches); and operation mode (1 D normally closed, 2 D normally
open). The lengths of observation periods are quite different, as indicated in the last column,
time. Using an offset for log of observation time, identify the factors associated with large numbers
of valve failures. (Data from L. M. Moore and R. J. Beckman, “Appropriate One-Sided Tolerance
Bounds on the Number of Failures Using Poisson Regression,” Technometrics 30 (1988): 283–90.)

```{r}
#read in the data
valves <- Sleuth3::ex2224

#take a look at the data
head(valves)

#scatterplot matrix of all the variables of interest 
ggpairs(valves, columns = (1:5))


# REGRESSION ANALYSIS USING BACKWARD ELIMINATION
values_glm1 <- glm(Failures ~ System + Operator + Valve + Size + Mode, family = poisson, data = valves)

```

<br>


<br>




