---
title: "Hankowsky_HW9"
output: html_document
date: "2022-11-03"
---

```{r setup, include=FALSE}
library(Sleuth3)
library(tidyverse)
library(asht)
library(mosaic)
library(GGally)
library(gridExtra)
library(olsrr)
library(car)
library(Hmisc)

```

Hankowsky Homework Solutions 


# Basic Problems 
## 18.10

```{r}
#create the table
polio <- matrix(c(142,199858,56,199944), nrow = 2, ncol = 2, byrow = TRUE,
                  dimnames = list(c("Placebo", "Salk Polio Vaccine"),c("Yes", "No")))

#look at the table
polio

# Part A 
fisher.test(polio, alternative="greater") # Fisher's exact test, 1-tailed


# Part B 
# Sampling distribution of a single population proportion 
#sample proportion 
142/198

#estimate the varinace of the sample population 
((142/198)*(1 - (142/198))) / (198)

#find the SE
sqrt(0.001024426)

#95% CI using normal multiplier 
0.7171717 + 0.03200666
0.7171717 - 0.03200666

#check endpoints for sample size requirements 
198*0.7491784
198*0.685165


#or just a one sample biomial tests
binom.test(142,198, p = 1/2, alternative = "greater")


# PRESENTATION PLOT 
#just looking at the data matrix
polio

#proportion tests
prop.test(142, (142+199858)) #Est = 0.00071; CI = 0.0006001657 to 0.0008394233
prop.test(56, (142+199944)) #Est = 0.0002798797; CI = 0.0002133819 to 0.0003662140

#saving the estimates and CIs from the prop tests
pHat   <- c(0.00071, 0.0002798797)       
lower95 <- c(0.0006001657, 0.0002133819)
upper95 <- c(0.0008394233, 0.0003662140)

polio_plot <- Cbind(pHat, lower95, upper95)

Treatment <- as.factor( c("Placebo", "Salk Polio Vaccine"))

#the plot 
Dotplot(Treatment ~ polio_plot, 
        xlab = "Probability of Infantile Paralysis Based on Treatment (and 95% Confidence Intervals)", 
        ylab = "Treatment", ylim=c(.5,2.5), cex=2)

```

<br>

These results provide overwhelming evidence that the odds of infantile paralysis are lower for those who received the Salk vaccine (one-sided p-value < 0.0001). The odds of infantile paralysis are 2.54 times greater for the placebo treatment than the Salk vaccine (95% CI: 1.94 to $\infty$). Yes, the observed proportion in the control group is drastically larger than would be excepted under the null hypothesis (one-sided p-value < 0.0001). 

<br>

***

## 19.23

Who Looks After the Kids? Exercises 10.31 and 17.13 considered data on clutch volume and
body mass to investigate the similarities of dinosaurs with several groups of currently living animals.
One issue concerning the validity of the study is the selection of the bird species in the set of currently
living animals. Were species (i) selected at random? (ii) stratified by group and selected within group
at random? (iii) selected based on what information was available? or (iv) selected to support the
study conclusions? Ignore the possibility of (iv). If the answer is (i) or (ii) then that would not cause
any problem with the analyses performed, but (iii) could potentially result in biased sampling and
therefore incorrect conclusions. Furthermore, (iii) seems like themost probable form of sampling of
the modern groups and almost certainly the form of sampling among the dinosaur groups. 

One way
to assess the likelihood (iii) for birds is to compare the numbers of species from each of the 29 orders
in the study with the known total number of species in each of the orders, as shown in Display 19.14.
If the selection of birds had been at random, the expected proportion of species in the study from one
particular order, n, is the proportion of all species in that order (N=9,866) times the total number of
species in the sample (414). That is, the expected number in each sample, if random sampling were
used, is .N=9,866/ ! 414. Calculate the expected numbers and compare the observed numbers with
them using Pearson’s chi-square statistic. It will have 29 " 1 D 28 d.f. What do you conclude?

```{r}
#Read-In the data
birds <- Sleuth3::ex1923

birds1 <- birds %>%
  mutate(expected_n = (N/9866)*414) %>%
  mutate(x = ((n-expected_n)^2)/expected_n) %>%
  mutate(x2 = sum(x))

chisq.test(birds1$expected_n, birds1$n)


fisher.test(birds1$expected_n, birds1$n)
```

<br>

***

# Supplemental Problems 
## 18.19

```{r}
#create the tables 
arrests <- matrix(c(28, 35, 36, 20), nrow = 2, ncol = 2, byrow = TRUE,
                  dimnames = list(c("No Preschool", "Preschool"),c("No", "Yes")))

income <- matrix(c(38, 25, 22, 34), nrow = 2, ncol = 2, byrow = TRUE,
                  dimnames = list(c("No Preschool", "Preschool"),c("No", "Yes")))

#look at the tables
arrests

income


#run the fishers test for arrests
fisher.test(arrests)

#take the inverse of the odds ratio to interpret it 
1/0.4475579 
#CIs
1/0.1983322 
1/0.9898925

#run the fishers test for income
fisher.test(income)



# PRESENTATION PLOT  for arrests
#just looking at the data matrix
arrests

#proportion tests
prop.test(35, (35+28)) #Est = 0.5555556  ; CI = 0.4255300   to 0.6787945
prop.test(20, (20+36)) #Est = 0.3571429  ; CI = 0.2368611   to 0.4970247

#saving the estimates and CIs from the prop tests
pHat   <- c(0.5555556 , 0.3571429)       
lower95 <- c(0.4255300 , 0.2368611)
upper95 <- c(0.6787945, 0.4970247)

arrests_plot <- Cbind(pHat, lower95, upper95)

#getting the matrix into the correct format
Treatment <- as.factor(c("No Preschool", "Preschool"))

#the plot 
p1 <- Dotplot(Treatment ~ arrests_plot, 
        xlab = "Probability of Getting Arrested 5 or More Times \n Based on Preschool Program (and 95% Confidence Intervals)", 
        ylab = "Treatment", ylim=c(.5,2.5), cex=2)




# PRESENTATION PLOT  for income
#just looking at the data matrix
income

#proportion tests
prop.test(25, (25+38)) #Est = 0.3968254   ; CI = 0.2782316   to 0.5280338
prop.test(34, (34+22)) #Est = 0.6071429   ; CI = 0.4675350    to 0.7320673

#saving the estimates and CIs from the prop tests
pHat   <- c(0.3968254 , 0.6071429)       
lower95 <- c(0.2782316 , 0.4675350)
upper95 <- c(0.5280338, 0.7320673)

arrests_plot2 <- Cbind(pHat, lower95, upper95)

#getting the matrix into the correct format
Treatment <- as.factor(c("No Preschool", "Preschool"))

#the plot 
p2 <- Dotplot(Treatment ~ arrests_plot2, 
        xlab = "Probability of Earning an Annual Income of $20K or More \n Based on Preschool Program (and 95% Confidence Intervals)", 
        ylab = "Treatment", ylim=c(.5,2.5), cex=2)

grid.arrange(p1, p2, ncol = 1)

```

<br>

These results provide moderate evidence that the odds of having 5 arrests or more are lower for those who participated in high-quality preschool programs (one-sided p-value = 0.042). The odds of having 5 arrests or more are 2.23 times less for participants in the high-quality preschool programs (95% CI: 1.01 to 5.04). These results provide solid evidence that the odds of having an annual income of $20K or more are greater for those who participated in high-quality preschool programs (one-sided p-value = 0.029). The odds of having an annual income 20K or more are 2.33 times greater for the participants in the high-quality preschool programs (95% CI: 1.06 to 5.24).

<br>

***

## 19.20

```{r}
#create the table
diet <- matrix(c(1, 44, 49, 228), nrow = 2, ncol = 2, byrow = TRUE,
                  dimnames = list(c("Women", "Men"),c("Dropped Out", "Completed")))

#look at the table
diet

#run the fishers test
fisher.test(diet)

#to get the odds ratio interpretative 
1/0.1061466 

#and the CIs 
1/0.002571065 
1/0.656355519


# PRESENTATION PLOT 
#just looking at the data matrix
diet

#proportion tests
prop.test(1, (1+44)) #Est = 0.02222222 ; CI = 0.001161097  to 0.132339248
prop.test(49, (49+228)) #Est = 0.1768953 ; CI = 0.1348649  to 0.2281157

#saving the estimates and CIs from the prop tests
pHat   <- c(0.02222222 , 0.1768953)       
lower95 <- c(0.001161097 , 0.1348649 )
upper95 <- c(0.132339248, 0.2281157)

diet_plot <- Cbind(pHat, lower95, upper95)

Gender <- as.factor( c("Women", "Men"))

#the plot 
Dotplot(Gender ~ diet_plot, 
        xlab = "Probability of Dropping out of the Diet Study (and 95% Confidence Intervals)", 
        ylab = "Treatment", ylim=c(.5,2.5), cex=2)

```

<br>

These data provide strong evidence that women were more likely to complete the diet study (p-value < 0.0063). The odds of completing the diet study were 9.42 times greater for woman than men (95% CI: 1.52 to 388.94). 

<br>
<br>
