---
title: "Hankowsky_HW9"
output: html_document
date: "2022-11-03"
---

<span style="color: magenta"> **General Comments:** Excellent work on all 4 problems! I have virtually no comments that would improve your work. **Basic score (5+5)/10, Supplemental Score (2.5+2.5)/5  Total score 15/10, Total Points: 145, Average (9 Homeworks): 161%,  Surplus: 65 points** </span>

```{r setup, include=FALSE}
library(Sleuth3)
library(tidyverse)
library(asht)
library(mosaic)
library(GGally)
library(gridExtra)
library(olsrr)
library(car)
library(Hmisc)

```

Hankowsky Homework Solutions 


# Basic Problems 
## 18.10

```{r}
#create the table
polio <- matrix(c(142,199858,56,199944), nrow = 2, ncol = 2, byrow = TRUE,
                  dimnames = list(c("Placebo", "Salk Polio Vaccine"),c("Yes", "No")))

#look at the table
polio

# Part A 
fisher.test(polio, alternative="greater") # Fisher's exact test, 1-tailed


# Part B 
# Sampling distribution of a single population proportion 
#sample proportion 
142/198

#estimate the varinace of the sample population 
((142/198)*(1 - (142/198))) / (198)

#find the SE
sqrt(0.001024426)

#95% CI using normal multiplier 
0.7171717 + 0.03200666
0.7171717 - 0.03200666

#check endpoints for sample size requirements 
198*0.7491784
198*0.685165


#or just a one sample biomial tests
binom.test(142,198, p = 1/2, alternative = "greater")


# PRESENTATION PLOT 
#just looking at the data matrix
polio

#proportion tests
prop.test(142, (142+199858)) #Est = 0.00071; CI = 0.0006001657 to 0.0008394233
prop.test(56, (142+199944)) #Est = 0.0002798797; CI = 0.0002133819 to 0.0003662140

#saving the estimates and CIs from the prop tests
pHat   <- c(0.00071, 0.0002798797)       
lower95 <- c(0.0006001657, 0.0002133819)
upper95 <- c(0.0008394233, 0.0003662140)

polio_plot <- Cbind(pHat, lower95, upper95)

Treatment <- as.factor( c("Placebo", "Salk Polio Vaccine"))

#the plot 
Dotplot(Treatment ~ polio_plot, 
        xlab = "Probability of Infantile Paralysis Based on Treatment (and 95% Confidence Intervals)", 
        ylab = "Treatment", ylim=c(.5,2.5), cex=2)

```

<br>

These results provide overwhelming evidence that the odds of infantile paralysis are lower for those who received the Salk vaccine (one-sided p-value < 0.0001). The odds of infantile paralysis are 2.54 times greater for the placebo treatment than the Salk vaccine (95% CI: 1.94 to $\infty$). Yes, the observed proportion in the control group is drastically larger than would be excepted under the null hypothesis (one-sided p-value < 0.0001). 

<span style="color: magenta"> **Basic 18.10** Superb work throughout. Chapter 18 hadn't yet introduced the Fisher test, but the Fisher test odds ratio (two tailed) will give you an estimate for the upper CI that isn't infinity **score 5/5** </span>

<br>

***

## 19.23
```{r}
#Read-In the data
birds <- Sleuth3::ex1923

birds <- birds %>%
  mutate(expected_n = (N/9866)*414) %>%
  mutate(expected_n2 = (N/sum(N))*sum(n)) %>%  #just making sure the calculation given by the textbook is the same as that given by prof Gallagher; it is 
  mutate(x = ((n-expected_n)^2)/expected_n) 

#just sorting the expected values
birds %>%
  arrange(expected_n)

#sum the expected values
chi2 <- sum(birds$x) 
  
#calculating the p-value for the chi-squared statistic with 28df, using 1-pchisq to get the area to the right of the curve
1-pchisq(chi2, 28)


```

<br>

This data provides overwhelming evidence that the selection of birds was not random (p-value < 0.00001). Thus, the selection of bird species is most likely based on the information that was available which could result in biased sampling and incorrect conclusions. 

<span style="color: magenta"> **Basic 19.23** Excellent work. You should never report a p value of 0 or 1. In my calcualtion of the chi2 it was pchisq(sum(birds$Chi2), df=28, lower.tail=FALSE), [1] 4.121023e-96 But, you can't report that either since machine precision for 0 is only 2e-16. Just report p<1e-6 **score 5/5** </span>

<br>

***

# Supplemental Problems 
## 18.19

```{r}
#create the tables 
arrests <- matrix(c(28, 35, 36, 20), nrow = 2, ncol = 2, byrow = TRUE,
                  dimnames = list(c("No Preschool", "Preschool"),c("No", "Yes")))

income <- matrix(c(38, 25, 22, 34), nrow = 2, ncol = 2, byrow = TRUE,
                  dimnames = list(c("No Preschool", "Preschool"),c("No", "Yes")))

#look at the tables
arrests

income


#run the fishers test for arrests
fisher.test(arrests)

#take the inverse of the odds ratio to interpret it 
1/0.4475579 
#CIs
1/0.1983322 
1/0.9898925

#run the fishers test for income
fisher.test(income)



# PRESENTATION PLOT  for arrests
#just looking at the data matrix
arrests

#proportion tests
prop.test(35, (35+28)) #Est = 0.5555556  ; CI = 0.4255300   to 0.6787945
prop.test(20, (20+36)) #Est = 0.3571429  ; CI = 0.2368611   to 0.4970247

#saving the estimates and CIs from the prop tests
pHat   <- c(0.5555556 , 0.3571429)       
lower95 <- c(0.4255300 , 0.2368611)
upper95 <- c(0.6787945, 0.4970247)

arrests_plot <- Cbind(pHat, lower95, upper95)

#getting the matrix into the correct format
Treatment <- as.factor(c("No Preschool", "Preschool"))

#the plot 
p1 <- Dotplot(Treatment ~ arrests_plot, 
        xlab = "Probability of Getting Arrested 5 or More Times \n Based on Preschool Program (and 95% Confidence Intervals)", 
        ylab = "Treatment", ylim=c(.5,2.5), cex=2)




# PRESENTATION PLOT  for income
#just looking at the data matrix
income

#proportion tests
prop.test(25, (25+38)) #Est = 0.3968254   ; CI = 0.2782316   to 0.5280338
prop.test(34, (34+22)) #Est = 0.6071429   ; CI = 0.4675350    to 0.7320673

#saving the estimates and CIs from the prop tests
pHat   <- c(0.3968254 , 0.6071429)       
lower95 <- c(0.2782316 , 0.4675350)
upper95 <- c(0.5280338, 0.7320673)

arrests_plot2 <- Cbind(pHat, lower95, upper95)

#getting the matrix into the correct format
Treatment <- as.factor(c("No Preschool", "Preschool"))

#the plot 
p2 <- Dotplot(Treatment ~ arrests_plot2, 
        xlab = "Probability of Earning an Annual Income of $20K or More \n Based on Preschool Program (and 95% Confidence Intervals)", 
        ylab = "Treatment", ylim=c(.5,2.5), cex=2)

grid.arrange(p1, p2, ncol = 1)

```

<br>

These results provide moderate evidence that the odds of having 5 arrests or more are lower for those who participated in high-quality preschool programs (one-sided p-value = 0.042). The odds of having 5 arrests or more are 2.23 times less for participants in the high-quality preschool programs (95% CI: 1.01 to 5.04). These results provide solid evidence that the odds of having an annual income of $20K or more are greater for those who participated in high-quality preschool programs (one-sided p-value = 0.029). The odds of having an annual income 20K or more are 2.33 times greater for the participants in the high-quality preschool programs (95% CI: 1.06 to 5.24).

<span style="color: magenta"> **Supplemental 18.19** Excellent analyses throughout and very nice graphics. **score 2.5/2.5** </span>

<br>

***

## 19.20

```{r}
#create the table
diet <- matrix(c(1, 44, 49, 228), nrow = 2, ncol = 2, byrow = TRUE,
                  dimnames = list(c("Women", "Men"),c("Dropped Out", "Completed")))

#look at the table
diet

#run the fishers test
fisher.test(diet)

#to get the odds ratio interpretative 
1/0.1061466 

#and the CIs 
1/0.002571065 
1/0.656355519


# PRESENTATION PLOT 
#just looking at the data matrix
diet

#proportion tests
prop.test(1, (1+44)) #Est = 0.02222222 ; CI = 0.001161097  to 0.132339248
prop.test(49, (49+228)) #Est = 0.1768953 ; CI = 0.1348649  to 0.2281157

#saving the estimates and CIs from the prop tests
pHat   <- c(0.02222222 , 0.1768953)       
lower95 <- c(0.001161097 , 0.1348649 )
upper95 <- c(0.132339248, 0.2281157)

diet_plot <- Cbind(pHat, lower95, upper95)

Gender <- as.factor( c("Women", "Men"))

#the plot 
Dotplot(Gender ~ diet_plot, 
        xlab = "Probability of Dropping out of the Diet Study (and 95% Confidence Intervals)", 
        ylab = "Treatment", ylim=c(.5,2.5), cex=2)

```

<br>

These data provide strong evidence that women were more likely to complete the diet study (p-value < 0.0063). The odds of completing the diet study were 9.42 times greater for woman than men (95% CI: 1.52 to 388.94). 

**Supplemental 19.20 I** Excellent throughout. You can get the odds ratios greater than 1 by just flipping a single row or column (not both) as shown in the following code chunk **Score 2.5/2.5** </span>

```{r Gallagher chunk}
myTable3 <- matrix(c(44,45-44,277-49,49), nrow = 2, ncol = 2, byrow = TRUE,
                   dimnames = list(c("Women", "Men"),
                                   c("Stayed", "Dropped")))
myTable3
fisher.test(myTable3)
fisher.test(myTable3, alternative="greater")
```

<br>
<br>
