---
title: "Hankowsky_HW10"
output: html_document
date: "2022-11-10"
---

<span style="color: magenta"> **General Comments:** Excellent work on the basic problems. **Basic score (5+5)/10, Supplemental Score (2.5+2.5)/5 Total score 15/10 Total Points (10 HW): , Average: , Surplus: ** </span>

```{r setup, include=FALSE}
library(Sleuth3)
library(tidyverse)
library(asht)
library(mosaic)
library(GGally)
library(gridExtra)
library(olsrr)
library(car)
library(Hmisc)
library(MASS)

```

Hankowsky Homework Solutions 


# Basic Problems 
## 20.16
```{r}
#read in the data
sparrows <- Sleuth3::ex2016

#take a look at the dataset 
head(sparrows)

#scatterplot matrix
ggpairs(sparrows, columns = 2:11, ggplot2::aes(color = Status))


#run the model 
#change to factor 
sparrows <- sparrows %>%
  mutate(AG = as.factor(AG))

sparrow_glm <- glm(Status ~ AG + TL + WT + BH + HL + FL + TT + SK + KL, family = binomial, data = sparrows)

crPlots(sparrow_glm)

#looks like there an effect of TL, WT, HL, and maybe a slight KL 
#after accounting for other variables

# REGRESSION ANALYSIS USING BACKWARD ELIMINATION
sparrow_glm1 <- glm(Status ~ AG + TL + WT + BH + HL + FL + TT + SK + KL, family = binomial, data = sparrows)
summary(sparrow_glm1)


sparrow_glm2 <- update(sparrow_glm1, ~ . - AG)
summary(sparrow_glm2)


sparrow_glm3 <- update(sparrow_glm2, ~ . - FL)
summary(sparrow_glm3)


sparrow_glm4 <- update(sparrow_glm3, ~ . - TT)
summary(sparrow_glm4)


sparrow_glm5 <- update(sparrow_glm4, ~ . - SK)
summary(sparrow_glm5)


sparrow_glm6 <- update(sparrow_glm5, ~ . - BH)
summary(sparrow_glm6)

#run the optimal model 
sparrow_glm_optimal <- glm(Status ~ TL + WT + HL + KL, family = binomial, data = sparrows)
summary(sparrow_glm_optimal)

#run the null model 
sparrow_glm_null <- glm(Status ~ 1, family = binomial, data = sparrows)
summary(sparrow_glm_null)

#drop in deviance 
anova(sparrow_glm_optimal, sparrow_glm_null, test = "Chisq")

```

<br>

These data provide overwhelming evidence that the probability of survival is associated with the physical characteristics
of the birds (Drop-in-Deviance Test, p-value < 0.00001). The physical characteristics of the birds that are most associated with the probability of survival are the total length from the tip of the beak to the tip of the tail in millimeters (p-value < 0.00001), weight in grams (p-value = 0.011), length of humerus in inches (p-value < 0.001), and the length of the keel of the sternum in inches (p-value = 0.020). 

<br>
<span style="color: magenta"> **Basic 20.16_3rd** Excellent work. You used a backward selection procedure to arrive at 4 key variables. If you check the VIFs, you'll find that with the full set of variables, many of the variables had VIFs greater than 4 indicating problems with multicollinearity. There are a variety of solutions that have been proposed: PC regression, Ridge regression, lasso, but the latest is to use Bayesian model averaging with Bayesian Information Content BIC as a criterion. I'll present that tonight, and you end up with the same 4 variables that you found: TL + WT + HL + KL **Score 5/5** </span>

***

## 21.18
```{r}
#read in the data
HIV <- Sleuth3::ex2118

#take a look at the dataset
head(HIV)

#create variables of interest 
Response1 <- cbind(ex2118$HIV, ex2118$Number-ex2118$HIV)
Circ <- ifelse(ex2118$Circumcised == "No", 1, 0)
Contact_S <- ifelse(ex2118$Contact == "Multiple", 1, 0)

#run glm 
HIV_glm <- glm(Response1 ~ Circ + Contact_S, family = binomial)

#get the odds ratio from the glm summary and back transform the odds ratio
summary(HIV_glm)
exp(cbind(OR = coef(HIV_glm), confint(HIV_glm)))

```

<br>

The odds of testing positive for HIV for people who were not circumcised was estimated to be 10.74 times the odds of people who were circumcised after controlling for the number of sexual contacts with prostitutes (95% CI: 4.25 to 31.01 times). The odds for testing positive for HIV for people who had multiple sexual contacts with prostitutes was estimated to be 1.45 times the odds of people who had a single sexual contact with a prostitute after controlling for the circumcision status (95% CI: 0.55 to 4.35 times). 

<span style="color: magenta"> **Basic 21.18_3rd** Excellent answer, but you should have noted that the evidence for an effect of multiple partners is weak (p = .48 ) and 95% CI which includes 1  **Score 5/5** </span>


<br>

***

# Supplemental Problems 
## 20.20
```{r}
#read in the data
esteem <- Sleuth3::ex1223

#take a look at the dataset
head(esteem)

#construct variables as illustrated in the problem 
esteem <- esteem %>%
  dplyr::select(Income2005, AFQT, Educ, Gender, Esteem1) %>%
  mutate(response_var = case_when(
    Esteem1 == 1 ~ 1, 
    Esteem1 %in% c(2,3,4) ~ 0)) %>%
  mutate(response_var = as.factor(response_var), 
         log_income = log(Income2005)) %>%
  relocate(response_var) %>%
  relocate(log_income, .after = Income2005)

#take a look at the new dataset
head(esteem)

#scatterplot matrix
ggpairs(esteem, columns = 3:5, ggplot2::aes(color = response_var))


#run the full model 
esteem_glm <- glm(response_var ~ log_income + AFQT + Educ + Gender, family = binomial, data = esteem)
crPlots(esteem_glm)


# REGRESSION ANALYSIS USING BACKWARD ELIMINATION
esteem_glm1 <- glm(response_var ~ log_income + AFQT + Educ + Gender, family = binomial, data = esteem)
summary(esteem_glm1)

esteem_glm2 <- update(esteem_glm1, ~ . - Gender)
summary(esteem_glm2)


#run the optimal model 
esteem_glm_optimal <- glm(response_var ~ log_income + AFQT + Educ, family = binomial, data = esteem)
summary(esteem_glm_optimal)

#back transform the estimates and CIs 
exp(cbind(OR = coef(esteem_glm_optimal), confint(esteem_glm_optimal)))

```

<br>

There is a 14.92% increase in the odds of having a positive self-esteem for every 10 fold increase in 2005 income, after controlling for AFQT score and years of education (95% CI: 5.38% to 25.36%). There is a 8.38% increase in the odds of having a positive self-esteem for every additional year of education, after controlling for income and AFQT score (95% CI: 4.04% to 12.93%). There is 1% increase in the odds of having a positive self-esteem for every one-unit increase in AFQT score, after controlling for income and years of education (95% CI: 0.04% to 1.1%)

<br>
<span style="color: magenta"> **Supplemental 20.20_3rd** Excellent answer!  Harrell (2015, 2021) argues that you shouldn't automatically drop explanatory variables just because the p value is greater than 0.05. You dropped gender, but you could have presented it too. Females have 14% higher self esteem than males (1-exp(-.146214)), but the 95% CI includes 1.    **Score 2.5/2.5** </span>

***

## 21.21

```{r}
#read in the data
wildflower <- Sleuth3::ex1221

#take a look at the dataset
head(wildflower)

#quick boxplot of the data 
wildflower %>%
  ggplot() + 
  geom_boxplot(aes(x = reorder(Rating, Total), y = Total)) + 
  labs(x = "Rating") + 
  theme_classic()

#fit an ordinail logistic regression model 
wildflower_polr <- polr(Rating ~ Total, data = wildflower)
summary(wildflower_polr)


#back transform the estimate 
exp(coef(wildflower_polr)) 
exp(confint(wildflower_polr))

```

<br>

For each unit increase in total rainfall in the months September to March, the wildflower display rating is only 79.82% as likely to increase rating status (95% CI: 71.76% to 88.15%)

<span style="color: magenta"> **Supplemental 21.21_3rd** You are just one of two students to tackle the ordinal logistic. I forgot to present the method in class. Your estimates of the effects are wrong, but I don't have time now to present the proper analysis. I'll provide an R chunk that produces a nice graphic of the response, which improves dramatically with rainfall. **Score 2.5/2.5** </span>

```{r Gallagher chunk}
library(car)
library(effects)  # for the Effect function, Fox & Weisberg (2019 p 274, 319)
library(Hmisc) # for describe function
library(MASS)
library(Sleuth3)
flower<-ex1221
str(flower)
describe(flower)
mod.polr <- polr(factor(Score) ~ Total, data = flower)
summary(mod.polr)
# Test for the proportional odds model and assumptions Fox & Weisberg
# (2019, p 322)
poTest(mod.polr)

# plot(Effect("Total"), mod.polr) # produces errors, don't know why

plot(Effect("Total", mod.polr),
     axes=list(y=list(style="stacked")),
     lines=list(col=gray(c(.1, .3, .5, .7, .9))),
     main = "ex21.21", xlab = "Total Rainfall (in) Sep - March")


```

<br>
<br>
