---
title: "Hankowsky_HW2"
output: html_document
date: "2022-09-14"
---

```{r setup, include=FALSE}
library(Sleuth3)
library(tidyverse)
library(Hmisc)
```

Hankowsky Homework Solutions 


# Basic Problems 

## 2.22
```{r}
#pulling the data from the 'Sleuth 3' package 
intelligence <- Sleuth3::ex0222

#transposing the data so it is in long format
intelligence <- intelligence %>%
  mutate(Gender = str_to_title(Gender)) %>%
  pivot_longer(cols = c(Arith, Word, Parag, Math), 
               names_to = "test", 
               values_to = "score") 

#boxplot of overall AFQT score by gender 
intelligence %>%
  ggplot() + 
  geom_boxplot(aes(x = Gender, y = AFQT, fill = Gender)) + 
  labs(title = "Boxplot of AFQT Score by Gender", 
    y = "Armed Forces Qualifying Test Score") + 
  theme_classic() 


#histogram of AFQT score by gender 
ggplot() + 
  geom_histogram(data = intelligence %>% 
                   filter(Gender == "Male"), 
                 aes(x = AFQT, y = ..density.., fill = Gender), color = "black") + 
  geom_histogram(data = intelligence %>% 
                   filter(Gender == "Female"), 
                 aes(x = AFQT, y = -..density.., fill = Gender), color = "black") + 
  labs(title = "Histogram of AFQT Score by Gender", 
    x = "Armed Forces Qualifying Test Score", y = "Denstiy") + 
  theme_classic()


#boxplot of individual test components by gender 
intelligence %>%
  ggplot() + 
  geom_boxplot(aes(x = test, y = score, fill = Gender)) +
  scale_x_discrete(labels = c("Arithmetic\nReasoning", "Mathematical\nKnowledge", "Paragraph\nComprehension", "Word\nKnowledge")) +  
  labs(title = "Boxplot of Individual Test Components of the AFQT by Gender", 
    x = "\nArmed Forces Qualifying Test Component", y = "Test Score") +
  theme_classic()


#histogram of individual test components by gender 
ggplot() + 
  geom_histogram(data = intelligence %>% 
                   filter(Gender == "Male"), 
                 aes(x = score, y = ..density.., fill = Gender), color = "black") + 
  geom_histogram(data = intelligence %>% 
                   filter(Gender == "Female"), 
                 aes(x = score, y = -..density.., fill = Gender), color = "black") + 
  labs(title = "Histogram of Individual Test Components of the AFQT by Gender", 
       x = "Test Score", y = "Density") + 
  theme_classic() +
  facet_wrap(~test, 
             labeller = as_labeller(c(
               Arith = "Arithmetic Reasoning", 
               Math = "Mathematical Knowledge", 
               Parag = "Paragraph Comprehension", 
               Word = "Word Knowledge")))



#running the t-test to see if there is a difference in AFQT score by gender 
#turning the gender variable into a factor
intel_t <- intelligence %>%
  mutate(Gender = as.factor(Gender))

#running the actual t-test
t.test(AFQT ~ Gender, var.equal = TRUE, data = intel_t)


#now looking at potential differences in each test component 
#creating a dataset for each test component
intel_t_arith <- intelligence %>%
  mutate(Gender = as.factor(Gender)) %>%
  filter(test == "Arith")

intel_t_word <- intelligence %>%
  mutate(Gender = as.factor(Gender)) %>%
  filter(test == "Word")

intel_t_parag <- intelligence %>%
  mutate(Gender = as.factor(Gender)) %>%
  filter(test == "Parag")

intel_t_math <- intelligence %>%
  mutate(Gender = as.factor(Gender)) %>%
  filter(test == "Math")


#running a t-test for each test component
#arith
t.test(score ~ Gender, var.equal = TRUE, data = intel_t_arith)

#word
t.test(score ~ Gender, var.equal = TRUE, data = intel_t_word)

#parag
t.test(score ~ Gender, var.equal = TRUE, data = intel_t_parag)

#math
t.test(score ~ Gender, var.equal = TRUE, data = intel_t_math)

```

<br>

The variance of the genders are about equal, so an equal variance two sample t-test was used. The histogram shows that both genders have a relatively uniform distributed, however, the two-sample t-test is robust to violations of normally distributed samples, so the analysis proceeded with an equal variance two sample t-test.These data provide strong evidence for a difference in male and female scoring on the overall Armed Forces Qualifying Test (two-sided p-value < 0.001 from an equal variance two-sample t-test). The mean score for males on the AFQT was 2.04 points greater than the mean female score (95% confidence interval: 0.97 to 3.11 points). 

The variance between gender was about equal for each of the individual components of the AFQT, so an equal variance two sample t-test was used. The histograms for the individual components of the test were again relatively uniform with the exception of the 'Paragraph Comprehension' section. However, the two-sample t-test is robust to violations of normally distributed samples, so the analysis again proceeded with an equal variance two sample t-test.There was overwhelming evidence that Arithmetic Reasoning section score differed by gender (p-value < 0.000001 from an equal variance two-sample t-test) with the mean score for males 2.04 points higher than the female score (95% confidence interval: 1.49 to 2.58 points). The Paragraph Comprehension section also had overwhelming evidence that the score differed by gender (p-value < 0.00001 from an equal variance two-sample t-test). The mean female score was 0.57 points greater than the mean male score (95% confidence interval 0.33 to 0.81 points). There was strong evidence for a difference in male and female test scores for the Mathematical Knowledge component as well (p-value = 0.0023 from an equal variance two-sample t-test) with the mean male score 0.75 points higher than the mean female score (95% confidence interval: 0.27 to 1.24 points). There was no evidence for a difference in male and female scoring on the Word Knowledge section (p-value = 0.94 from an equal variance two-sample t-test). 



***

## 3.32

#### A
```{r}
#pulling the data from the 'Sleuth 3' package
tuition <- Sleuth3::ex0332

#creating a column with the difference of outofstate tuition minus instate tuition
tuition <- tuition %>%
  mutate(diff = OutOfState - InState)

#transposing the data so it's in long format 
tuition_long <- tuition %>%
  pivot_longer(cols = c(InState, OutOfState), 
               names_to = "State", 
               values_to = "Tuition")  %>%
  mutate(log_tuition = log(Tuition))


#boxplot of in-state v out-of-state tuition for public universities
tuition_long %>%
  filter( Type == "Public") %>%
  ggplot() + 
  geom_boxplot(aes(x = State, y = Tuition, fill = State)) + 
  scale_x_discrete(labels = c("In-State", "Out-Of-State")) +
  scale_fill_discrete(name = "", labels = c("In-State", "Out-Of-State")) +
  labs(title = "Boxplot of In-State vs Out-Of-State Tuition for Public Universities", 
       x = "") + 
  theme_classic()


#histogram of in-state v out-of-state tuition for public universities
tuition_long %>%
  filter( Type == "Public") %>%
  ggplot() + 
  geom_histogram(data = tuition_long %>% 
                   filter(State == "InState"), 
                 aes(x = Tuition, y = ..density.., fill = State), color = "black") + 
  geom_histogram(data = tuition_long %>% 
                   filter(State == "OutOfState"), 
                 aes(x = Tuition, y = -..density.., fill = State), color = "black") + 
  scale_fill_discrete(name = "", labels = c("In-State", "Out-Of-State")) +
  labs(title = "Histogram of In-State vs Out-Of-State Tuition for Public Universities", 
       y = "Density") +
  theme_classic()

#boxplot of the difference in tuitions
tuition %>%
  filter(Type == "Public") %>%
  ggplot() +
  geom_boxplot(aes(y = diff)) +
  labs(title = "Tuition Difference: Out-Of-State minus In-State for Public Universities",
       y = "Difference in Out-Of-State and In-State Tution ($)",
       x = "50 Colleges") +
  theme_classic()

#histogram of the difference in tuitions
tuition %>%
  filter(Type == "Public") %>%
  ggplot() +
  geom_histogram(aes(x = diff)) +
  labs(title = "Tuition Difference: Out-Of-State minus In-State for Public Universities",
       x = "Difference in Out-Of-State and In-State Tution ($)",
       y = "Count") +
  theme_classic()


#log transforming the data and looking at the difference of logs 
tuition <- tuition %>%
  mutate(log_outofstate = log(OutOfState),
         log_instate = log(InState),
         diff_of_logs = log_outofstate - log_instate)



#boxplot of in-state v out-of-state tuition for public universities
tuition_long %>%
  filter( Type == "Public") %>%
  ggplot() + 
  geom_boxplot(aes(x = State, y = log_tuition, fill = State)) + 
  scale_x_discrete(labels = c("In-State", "Out-Of-State")) +
  scale_fill_discrete(name = "", labels = c("In-State", "Out-Of-State")) +
  labs(title = "Boxplot of Ln(In-State vs Out-Of-State Tuition) for Public Universities", 
       x = "", y = "Ln(Tuition)") + 
  theme_classic()


#histogram of in-state v out-of-state tuition for public universities
tuition_long %>%
  filter( Type == "Public") %>%
  ggplot() + 
  geom_histogram(data = tuition_long %>% 
                   filter(State == "InState"), 
                 aes(x = log_tuition, y = ..density.., fill = State), color = "black") + 
  geom_histogram(data = tuition_long %>% 
                   filter(State == "OutOfState"), 
                 aes(x = log_tuition, y = -..density.., fill = State), color = "black") + 
  scale_fill_discrete(name = "", labels = c("In-State", "Out-Of-State")) +
  labs(title = "Histogram of Ln(In-State vs Out-Of-State Tuition) for Public Universities", 
       y = "Density", x = "Ln(Tuition)") +
  theme_classic()




#boxplot of the difference in logged tuition
tuition %>%
  filter(Type == "Public") %>%
  ggplot() +
  geom_boxplot(aes(y = diff_of_logs)) + 
  labs(title = "Ln(Tuition Difference: Out-Of-State minus In-State) for Public Universities",
       y = "Ln(Difference in Out-Of-State and In-State Tution ($))",
       x = "50 Colleges") +
  theme_classic()

#histogram of the difference in logged tuition
tuition %>%
  filter(Type == "Public") %>%
  ggplot() +
  geom_histogram(aes(x = diff_of_logs)) + 
  labs(title = "Ln(Tuition Difference: Out-Of-State minus In-State) for Public Universities",
       x = "Ln(Difference in Out-Of-State and In-State Tution ($))",
       y = "Count") +
  theme_classic()




#t-test time
#in/out of state tuition in public
tuition_public_t <- tuition_long %>%
  filter(Type == "Public") %>%
  mutate(State = as.factor(State))

t1 <- t.test(log_tuition ~ State, data = tuition_public_t, paired = TRUE)
t1

#back transform estimate and CI
exp(t1$est[1])
exp(t1$conf.int)

```

<br>

The histograms for the data show that the in-state and out-of-state data for public universities are not normally distributed. However, because the data are paired, two observations (in-state/out-of-state tuition) from one study unit (university) the difference in tuition needs to be investigated. The difference in tuitions is not normally distributed and is not symmetric, so a log-transformation was performed. The histograms of the log-transformed data show that they are relatively normally distributed, so the analysis was conducted with the log-transformed data. The The difference in tuitions is not normally distributed but the analysis was conducted anyway, because t-tests are robust to deviations from normality. These data provide substantial evidence that the mean difference between in-state and out-of-state tuitions in public universities in nonzero (two-sided p-value < 0.00001 from a paired paired t-test). The mean for in-state tuition for private universities was estimated to exceed the mean for in-state tuition for public universities by 0.44 dollars (95% confidence interval: 0.38 to 0.50 dollars).



#### B 
```{r}
#boxplot of in-state tuition by university type 
tuition %>%
  ggplot() + 
  geom_boxplot(aes(x = Type, y = InState, fill = Type)) + 
  labs(title = "Boxplot of In-State Tuition by University Type", 
       x = "Univeristy Type", y = "In-State Tuition ($)") + 
  theme_classic()


#histogram of in-state tuition by university type 
ggplot() + 
  geom_histogram(data = tuition %>% 
                   filter(Type == "Public"), 
                 aes(x = InState, y = ..density.., fill = Type), color = "black") + 
  geom_histogram(data = tuition %>% 
                   filter(Type == "Private"), 
                 aes(x = InState, y = -..density.., fill = Type), color = "black") + 
  labs(title = "Histogram of In-State Tuition by University Type", 
       y = "Density", x = "In-State Tuition ($)") +
  theme_classic()





#boxplot of Ln(in-state tuition) by university type 
tuition %>%
  ggplot() + 
  geom_boxplot(aes(x = Type, y = log_instate, fill = Type)) + 
  labs(title = "Boxplot of Ln(In-State Tuition) by University Type", 
       x = "Univeristy Type", y = "Ln(In-State Tuition ($))") + 
  theme_classic()


#histogram of Ln(in-state tuition) by university type 
ggplot() + 
  geom_histogram(data = tuition %>% 
                   filter(Type == "Public"), 
                 aes(x = log_instate, y = ..density.., fill = Type), color = "black") + 
  geom_histogram(data = tuition %>% 
                   filter(Type == "Private"), 
                 aes(x = log_instate, y = -..density.., fill = Type), color = "black") + 
  labs(title = "Histogram of Ln(In-State Tuition) by University Type", 
       y = "Density", x = "Ln(In-State Tuition ($))") +
  theme_classic()



#t-test
#public v private in state tuition 
t2 <- t.test(log_instate ~ Type, var.equal = FALSE, data = tuition)
t2

#back transform estimate and CI
exp(t2$est[1]) - exp(t2$est[2])
exp(t2$conf.int)

```

<br>

The histograms show that the in-state tuition data are not normally distributed, so a log-transformation was performed. The histograms of the log-transformed data show that they are relatively normally distributed, so the analysis was conducted with the log-transformed data. The variance for the public and private universities are not equal, so an unequal variance two sample t-test was used. These data provide strong evidence for a difference in in-state tuition between public and private universities (two-sided p-value < 0.00001 from an unequal variance two-sample t-test). The mean for in-state tuition for private universities was estimated to exceed the mean for in-state tuition for public universities by $18,199 (95% confidence interval: 2.90 to 4.70 dollars).


#### C 
```{r}
#boxplot of out-of-state tuition by university type 
tuition %>%
  ggplot() + 
  geom_boxplot(aes(x = Type, y = OutOfState, fill = Type)) + 
  labs(title = "Boxplot of Out-Of-State Tuition by University Type", 
       x = "Univeristy Type", y = "Out-Of-State Tuition ($)") + 
  theme_classic()


#histogram of out-of-state tuition by university type 
ggplot() + 
  geom_histogram(data = tuition %>% 
                   filter(Type == "Public"), 
                 aes(x = OutOfState, y = ..density.., fill = Type), color = "black") + 
  geom_histogram(data = tuition %>% 
                   filter(Type == "Private"), 
                 aes(x = OutOfState, y = -..density.., fill = Type), color = "black") + 
  labs(title = "Histogram of Out-Of-State Tuition by University Type", 
       y = "Density", x = "Out-Of-State Tuition ($)") +
  theme_classic()





#boxplot of Ln(out-of-state tuition) by university type 
tuition %>%
  ggplot() + 
  geom_boxplot(aes(x = Type, y = log_outofstate, fill = Type)) + 
  labs(title = "Boxplot of Ln(Out-Of-State Tuition) by University Type", 
       x = "Univeristy Type", y = "Ln(Out-Of-State Tuition ($))") + 
  theme_classic()


#histogram of Ln(out-of-state tuition) by university type 
ggplot() + 
  geom_histogram(data = tuition %>% 
                   filter(Type == "Public"), 
                 aes(x = log_outofstate, y = ..density.., fill = Type), color = "black") + 
  geom_histogram(data = tuition %>% 
                   filter(Type == "Private"), 
                 aes(x = log_outofstate, y = -..density.., fill = Type), color = "black") + 
  labs(title = "Histogram of Ln(Out-Of-State Tuition) by University Type", 
       y = "Density", x = "Ln(Out-Of-State Tuition ($))") +
  theme_classic()




#t-test 
#public v private out of state tuition 
t3 <- t.test(log_outofstate ~ Type, var.equal = FALSE, data = tuition)
t3

#back transform estimate and CI
exp(t3$est[1]) - exp(t3$est[2])
exp(t3$conf.int)
```

<br>

The histograms show that the out-of-state tuition data are not normally distributed, so a log-transformation was performed. The histograms of the log-transformed data show that they are relatively normally distributed, so the analysis was conducted with the log-transformed data. The variance for the public and private universities are not equal, so an unequal variance two sample t-test was used. These data provide strong evidence for a difference in out-of-state tuition between public and private universities (two-sided p-value < 0.001 from an unequal variance two-sample t-test). The mean for out-of-state tuition for private universities was estimated to exceed the mean for out-of-state tuition for public universities by $9,488 (95% confidence interval: 1.26 to 2.05 dollars).


***

# Supplemental Problems 

## 2.21 (Sleuth 2nd Edition)
```{r}
#pulling the data from the 'Sleuth 2' package 
birds <- Sleuth2::ex0221

#just cleaning up the dataset to make it easier for plotting; capitalizing the status
birds <- birds %>%
  mutate(Status = str_to_title(Status))


#boxplot of bird weight and survival
birds %>%
  ggplot() + 
  geom_boxplot(aes(x = Status, y = Weight, fill = Status)) + 
  labs (title = "Boxplot of Bird Weight and Survival", 
    x = "", y = "Weight (g)") + 
  theme_classic()


#histogram of bird weight and survival 
ggplot() + 
  geom_histogram(data = birds %>% 
                   filter(Status == "Survived"), 
                 aes(x = Weight, y = ..density.., fill = Status), color = "black") + 
  geom_histogram(data = birds %>% 
                   filter(Status == "Perished"), 
                 aes(x = Weight, y = -..density.., fill = Status), color = "black") + 
  labs (title = "Histogram of Bird Weight and Survival", 
    x = "Weight (g)", y = "Density") + 
  theme_classic()



#running the t-test to see if there is a difference in bird survival by weight 
#turning the survival variable into a factor
birds_t <- birds %>%
  mutate(Status = as.factor(Status))

#running the actual t-test
t.test(Weight ~ Status, var.equal = FALSE, data = birds_t) 

```

<br>


The variance for the bird status were not equal, so an unequal variance two sample t-test was used. The histograms show that both the survival and perished groups are relatively normally distributed. These data provide moderate evidence for a difference in survival based on the house sparrow's weight (two-sided p-value = 0.033 from an unequal variance two-sample t-test). The mean for the birds that perished was estimated to exceed the mean for the birds that survived by 0.81 grams (95% confidence interval: 0.70 to 1.55 grams).



***

## 3.31 (Sleuth 2nd Edition)
```{r}
#pulling the data from the 'Sleuth 2' package 
mice <- Sleuth2::ex0331

#boxplot of mice iron retained and supplement 
mice %>%
  ggplot() + 
  geom_boxplot(aes(x = Supplement, y = Iron, fill = Supplement)) + 
  scale_x_discrete(labels = c(expression(Fe^"3+"), expression(Fe^"4+"))) +
  scale_fill_discrete(labels = c(expression(Fe^"3+"), expression(Fe^"4+"))) +
  labs (title = "Boxplot of Percent Iron Retained in Mice by Iron Supplement", 
    x = "Iron Supplement", y = "Iron Retained (%)") + 
  theme_classic()


#histogram of mice iron retained and supplement 
ggplot() + 
  geom_histogram(data = mice %>% 
                   filter(Supplement == "Fe3"), 
                 aes(x = Iron, y = ..density.., fill = Supplement), color = "black") + 
  geom_histogram(data = mice %>% 
                   filter(Supplement == "Fe4"), 
                 aes(x = Iron, y = -..density.., fill = Supplement), color = "black") + 
  scale_fill_discrete(labels = c(expression(Fe^"3+"), expression(Fe^"4+"))) +
  labs (title = "Histogram of Percent Iron Retained in Mice by Iron Supplement", 
    x = "Iron Retained (%)", y = "Density") + 
  theme_classic()





#log transforming the data 
mice <- mice %>%
  mutate(log_iron = log(Iron))


#boxplot of Ln(iron retained) and supplement 
mice %>%
  ggplot() + 
  geom_boxplot(aes(x = Supplement, y = log_iron, fill = Supplement)) + 
  scale_x_discrete(labels = c(expression(Fe^"3+"), expression(Fe^"4+"))) +
  scale_fill_discrete(labels = c(expression(Fe^"3+"), expression(Fe^"4+"))) +
  labs (title = "Boxplot of Ln(Percent Iron Retained) in Mice by Iron Supplement", 
    x = "Iron Supplement", y = "Ln(Iron Retained (%))") + 
  theme_classic()


#histogram of Ln(iron retained) and supplement 
ggplot() + 
  geom_histogram(data = mice %>% 
                   filter(Supplement == "Fe3"), 
                 aes(x = log_iron, y = ..density.., fill = Supplement), color = "black") + 
  geom_histogram(data = mice %>% 
                   filter(Supplement == "Fe4"), 
                 aes(x = log_iron, y = -..density.., fill = Supplement), color = "black") + 
  scale_fill_discrete(labels = c(expression(Fe^"3+"), expression(Fe^"4+"))) +
  labs (title = "Histogram of Ln(Percent Iron Retained) in Mice by Iron Supplement", 
    x = "Ln(Iron Retained (%))", y = "Density") + 
  theme_classic()




#running the t-test to see if there is a difference in mouse percentage of iron retained by iron supplement type 
#turning the survival variable into a factor
mice_t <- mice %>%
  mutate(Supplement = as.factor(Supplement))

#running the actual t-test
t1 <- t.test(log_iron ~ Supplement, var.equal = FALSE, data = mice_t)
t1

#back transform estimate and CI
exp(t1$est[2])/exp(t1$est[1])
exp(-t1$conf.int)




```

<br>


The histograms show that the iron supplement groups are not normally distributed, so a log-transformation was performed. The histograms of the log-transformed data show that they are normally distributed, so the analysis was conducted with the log-transformed data. The variance for the two iron supplement groups are not equal, so an unequal variance two sample t-test was used. These data provide moderate evidence for a difference in percent iron retained based on the iron supplement type (two-sided p-value = 0.0059 from an unequal variance two-sample t-test). The mean for the Fe4+ iron supplement was 1.68 percent greater than the mean for the Fe3+ iron supplement group (95% confidence interval: 1.17 to 2.41 %). 

<br>
<br>
