---
title: "Final_Project"
output: html_document
date: "2022-11-14"
---

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(skimr)
library(data.table)
library(fuzzyjoin)
library(lubridate)
library(kableExtra)
library(modEvA)
library(GGally)

library(maps) # land masses
library(rgdal) # read OGR vector maps into Spatial objects
library(viridis) # for customizing scale bars
library(mapproj)
library(hexbin)
library(sf)
library(scales)
library(gridExtra)
library(openxlsx)
library(beepr)
library(purrr)
library(rsq)
library(car)
library(jtools)
```

# Read- In the Data
```{r}
paired_tows_1mi_24hr <- read.csv('./data/paired_tows_1mi_24hr.csv')
paired_tows_5mi_72hr <- read.csv('./data/paired_tows_5mi_72hr.csv')

paired_tows_1mi_24hr_ob <- read.csv('./data/paired_tows_1mi_24hr_ob.csv')
paired_tows_5mi_72hr_ob <- read.csv('./data/paired_tows_5mi_72hr_ob.csv')
```

# Data for Maps 
```{r}
#Additional Data for Maps
#Statistical Areas
sareas <- st_read(dsn = "./data/Statistical_Areas_2010_withNames") %>%
  st_set_crs(4269)
sareas <- sareas %>% filter(Id %in% c("464", "465", "511", "512", "513", "514", "515", "521", "522", "525", "526", "537", "538", "539", "551", "561", "562", "611", "612", "613", "543", "542", "541", "534", "533", "616", "615", "614", "621", "622", "623", "624"))  
centroids <- sf::st_coordinates(st_centroid(sareas)) 
areas <- cbind(sareas, centroids)


#Land Mass
states <- map_data("state")
NEUS <- subset(states, region %in% c("Maine", "massachusetts", "new hampshire", "vermont", "maine", "rhode island",
                "connecticut","new york","new jersey","delaware","pennsylvania",
                            "maryland","virginia","north carolina","west virginia"))


#NEFSC Survey Strata
sf::sf_use_s2(FALSE)
nefsc_strata <- st_read(dsn = "./data/strata") %>%
  st_set_crs(4269)


```

# Function Creation 
```{r}
##############################################################################
#################### FUNCTIONS FOR T-TESTS ###################################
##############################################################################

#plots to check for normality before t-tests 
#boxplot of catch per unit area
box.plot <- function(x) {
  x %>% 
  pivot_longer(cols = c(sf_species_cpua, nefsc_species_cpua), 
               names_to = "Trawl", 
               values_to = "CPUA") %>%
  filter(CPUA < 10000) %>%
  ggplot() + 
  geom_boxplot(aes(x = Trawl, y = CPUA)) + 
  labs(x = "Trawl Type") + 
  scale_x_discrete(labels = c("nefsc_species_cpua" = "NEFSC", 
                              "sf_species_cpua" = "Study Fleet")) + 
  theme_classic()
}

#histogram of catch per unit area
hist_kh <- function(x) {
  x %>%  
  ggplot() + 
  geom_histogram(aes(x = sf_species_cpua, y = ..density..), fill = "red") + 
  geom_label(aes(x = Inf, y = Inf, label = "Study Fleet CPUE", vjust =1 , hjust = 1), color = "red") +
  geom_histogram(aes(x = nefsc_species_cpua, y = -..density..), fill = "blue") + 
  geom_label(aes(x = Inf, y = -Inf, label = "NEFSC CPUE", vjust = -1, hjust = 1), color = "blue") +
  labs(x = "CPUA (lb/min)", y = "Density", 
       title = "Histogram of CPUE for x", 
       subtitle = "Pairs within 5 mi and 72 hrs") + 
  xlim(0,2000) + 
  theme_classic()
}
  

#boxplot with log-transformed data 
box.log.plot <- function(x) {
  x %>% 
  pivot_longer(cols = c(log_sf_species_cpua, log_nefsc_species_cpua), 
               names_to = "Trawl", 
               values_to = "CPUA") %>%
  filter(CPUA < 10000) %>%
  ggplot() + 
  geom_boxplot(aes(x = Trawl, y = CPUA)) + 
  labs(x = "Trawl Type", y = "Log(CPUA)") + 
  scale_x_discrete(labels = c("log_nefsc_species_cpua" = "NEFSC", 
                              "log_sf_species_cpua" = "Study Fleet")) + 
  theme_classic()
}

#histogram with log-transformed data 
hist_log_kh <- function(x) {
  x %>%  
  ggplot() + 
  geom_histogram(aes(x = log_sf_species_cpua, y = ..density..), fill = "red") + 
  # geom_label(aes(x = Inf, y = Inf, label = "Study Fleet CPUE", vjust =1 , hjust = 1), color = "red") +
  geom_histogram(aes(x = log_nefsc_species_cpua, y = -..density..), fill = "blue") + 
  # geom_label(aes(x = Inf, y = -Inf, label = "NEFSC CPUE", vjust = -1, hjust = 1), color = "blue") +
  labs(x = "Log(CPUA (lb/min))", y = "Density", 
       title = "Histogram of CPUE for x", 
       subtitle = "Pairs within 5 mi and 72 hrs") + 
  theme_classic()
}


############## FUNCTIONS FOR T-TESTS FOR OBSERVER DATA #######################

#plots to check for normality before t-tests 
#boxplot of catch per unit area
box.plot_ob <- function(x) {
  x %>% 
  pivot_longer(cols = c(ob_species_cpua, nefsc_species_cpua), 
               names_to = "Trawl", 
               values_to = "CPUA") %>%
  filter(CPUA < 10000) %>%
  ggplot() + 
  geom_boxplot(aes(x = Trawl, y = CPUA)) + 
  labs(x = "Trawl Type") + 
  scale_x_discrete(labels = c("nefsc_species_cpua" = "NEFSC", 
                              "ob_species_cpua" = "Observer")) + 
  theme_classic()
}

#histogram of catch per unit area
hist_kh_ob <- function(x) {
  x %>%  
  ggplot() + 
  geom_histogram(aes(x = ob_species_cpua, y = ..density..), fill = "red") + 
  geom_label(aes(x = Inf, y = Inf, label = "Observer CPUE", vjust =1 , hjust = 1), color = "red") +
  geom_histogram(aes(x = nefsc_species_cpua, y = -..density..), fill = "blue") + 
  geom_label(aes(x = Inf, y = -Inf, label = "NEFSC CPUE", vjust = -1, hjust = 1), color = "blue") +
  labs(x = "CPUA (lb/min)", y = "Density", 
       title = "Histogram of CPUE for x", 
       subtitle = "Pairs within 5 mi and 72 hrs") + 
  xlim(0,2000) + 
  theme_classic()
}
  

#boxplot with log-transformed data 
box.log.plot_ob <- function(x) {
  x %>% 
  pivot_longer(cols = c(log_ob_species_cpua, log_nefsc_species_cpua), 
               names_to = "Trawl", 
               values_to = "CPUA") %>%
  filter(CPUA < 10000) %>%
  ggplot() + 
  geom_boxplot(aes(x = Trawl, y = CPUA)) + 
  labs(x = "Trawl Type", y = "Log(CPUA)") + 
  scale_x_discrete(labels = c("log_nefsc_species_cpua" = "NEFSC", 
                              "log_ob_species_cpua" = "Observer")) + 
  theme_classic()
}

#histogram with log-transformed data 
hist_log_kh_ob <- function(x) {
  x %>%  
  ggplot() + 
  geom_histogram(aes(x = log_ob_species_cpua, y = ..density..), fill = "red") + 
  # geom_label(aes(x = Inf, y = Inf, label = "Study Fleet CPUE", vjust =1 , hjust = 1), color = "red") +
  geom_histogram(aes(x = log_nefsc_species_cpua, y = -..density..), fill = "blue") + 
  # geom_label(aes(x = Inf, y = -Inf, label = "NEFSC CPUE", vjust = -1, hjust = 1), color = "blue") +
  labs(x = "Log(CPUA (lb/min))", y = "Density", 
       title = "Histogram of CPUE for x", 
       subtitle = "Pairs within 5 mi and 72 hrs") + 
  theme_classic()
}



##############################################################################
######################## FUNCTIONS FOR MODELLING #############################
##############################################################################

#trying to  create functions because I will be running the same models for every species dataset 
covars <- c("season", "mgmt_area", "AVGDEPTH", "BOTTEMP")
combos <- purrr::map(1:4, ~as.data.frame(t(combn(covars, .x))))

formulae <- combos %>%
  map_dfr(unite, col="formula", sep = " + ") %>%
  mutate(formula = paste0("relative_efficiency ~  VESSEL_NAME +", formula))


#fit models (takes dataset)
fit.models <- function(x) {
  formulae$formula %>% purrr::map( ~glm(.x, family = binomial, data=x))
}


#compare models (takes list of models created by fit.models())
compare.models <- function(x) {
  formulae %>%
  mutate(aic =  purrr::map_dbl(x, AIC)) %>%
  arrange(desc(aic)) %>%
  mutate(deltaAIC = aic - aic[15]) %>%
  mutate(rsq =  purrr::map_dbl(x, Dsquared)) %>% 
  mutate(rsq = sprintf("%0.2f", rsq)) %>%
  mutate(aic = sprintf("%1.0f", aic)) %>%
  mutate(deltaAIC = sprintf("%1.0f", deltaAIC)) %>%
  kbl(col.names = (c("Model", "AIC", "deltaAIC","Deviance Explained"))) %>%
  kable_classic("striped", full_width = FALSE)
}



############## FUNCTIONS FOR MODELING OBSERVER DATA ###########################
#trying to  create functions because I will be running the same models for every species dataset 
covars <- c("season", "mgmt_area", "AVGDEPTH", "BOTTEMP")
combos <- purrr::map(1:4, ~as.data.frame(t(combn(covars, .x))))

formulae_ob <- combos %>%
  map_dfr(unite, col="formula", sep = " + ") %>%
  mutate(formula = paste0("relative_efficiency ~  trawl_type +", formula))


#fit models (takes dataset)
fit.models_ob <- function(x) {
  formulae_ob$formula %>% purrr::map( ~glm(.x, family = binomial, data=x))
}


#compare models (takes list of models created by fit.models())
compare.models_ob <- function(x) {
  formulae_ob %>%
  mutate(aic =  purrr::map_dbl(x, AIC)) %>%
  arrange(desc(aic)) %>%
  mutate(deltaAIC = aic - aic[15]) %>%
  mutate(rsq =  purrr::map_dbl(x, Dsquared)) %>% 
  mutate(rsq = sprintf("%0.2f", rsq)) %>%
  mutate(aic = sprintf("%1.0f", aic)) %>%
  mutate(deltaAIC = sprintf("%1.0f", deltaAIC)) %>%
  kbl(col.names = (c("Model", "AIC", "deltaAIC","Deviance Explained"))) %>%
  kable_classic("striped", full_width = FALSE)
}

```



# Organizing the Paired Tow Data
```{r}
#filtering for tows cod and yellowtail were caught in both nets and calculating relative efficiency in the study fleet dataset 
paired_tows_5mi_72hr_re <- paired_tows_5mi_72hr %>%
  filter(SVSPP == 73 & SPECIES_ITIS == 164712 | #cod 
         SVSPP == 105 & SPECIES_ITIS == 172909  #yellowtail flounder
         ) %>%
  drop_na(sf_species_cpua, nefsc_species_cpua) %>%
  filter(CATCHWT_updated > 0) %>%
  filter(sf_species_cpua > 0) %>%
  mutate(relative_efficiency = sf_species_cpua/(nefsc_species_cpua + sf_species_cpua)) %>%
  mutate(mgmt_area = case_when(
    AREA_CODE %in% c(511, 512, 513, 514, 515) ~ "GOM", 
    AREA_CODE %in% c(521, 522, 526, 525, 561, 562, 551, 552) ~ "GB", 
    AREA_CODE %in% c(611, 612, 613, 539, 538, 537) ~ "SNE"
  )) %>%
  mutate(season = case_when(
    month(START_TOW_DATE_GMT.x) %in% c(03, 04, 05) ~ "Spring", 
    month(START_TOW_DATE_GMT.x) %in% c(09, 10, 11) ~ "Fall", 
  )) %>%
  distinct() 


#filtering for tows cod and yellowtail were caught in both nets and calculating relative efficiency in the observer dataset 
paired_tows_5mi_72hr_ob_re <- paired_tows_5mi_72hr_ob %>%
  filter(SVSPP == 73 & NESPP4 == 818 | #cod 
         SVSPP == 105 & NESPP4 == 1230 #yellowtail flounder
         ) %>%
  drop_na(ob_species_cpua, nefsc_species_cpua) %>%
  filter(CATCHWT_updated > 0) %>%
  mutate(relative_efficiency = ob_species_cpua/(nefsc_species_cpua + ob_species_cpua)) %>%
  mutate(mgmt_area = case_when(
    AREA.x %in% c(511, 512, 513, 514, 515) ~ "GOM", 
    AREA.x %in% c(521, 522, 526, 525, 561, 562, 551, 552) ~ "GB", 
    AREA.x %in% c(611, 612, 613, 539, 538, 537) ~ "SNE"
  )) %>%
  mutate(season = case_when(
    month(START_TOW_DATE_GMT) %in% c(03, 04, 05) ~ "Spring", 
    month(START_TOW_DATE_GMT) %in% c(09, 10, 11) ~ "Fall", 
  )) %>%
  mutate(trawl_type = as.factor(trawl_type)) %>%
  distinct()
```

# Species Datasets 
```{r}
#creating a separate dataset for each species 
#STUDY FLEET
#COD
cod_paired_tows_5mi_72hr_re <- paired_tows_5mi_72hr_re %>%
  filter(SVSPP == 73 & SPECIES_ITIS == 164712) %>% 
  distinct(EFFORT_ID, .keep_all = TRUE)

#YELLLOWTAIL FLOUNDER
yellowtailfl_paired_tows_5mi_72hr_re <- paired_tows_5mi_72hr_re %>%
  filter(SVSPP == 105 & SPECIES_ITIS == 172909) %>% 
  distinct(EFFORT_ID, .keep_all = TRUE)



#OBSERVER
#COD
cod_paired_tows_5mi_72hr_ob_re <- paired_tows_5mi_72hr_ob_re %>%
  filter(SVSPP == 73 & NESPP4 == 818) %>% 
  distinct(trip_haul_ID, .keep_all = TRUE)

#YELLLOWTAIL FLOUNDER
yellowtailfl_paired_tows_5mi_72hr_ob_re <- paired_tows_5mi_72hr_ob_re %>%
  filter(SVSPP == 105 & NESPP4 == 1230) %>% 
  distinct(trip_haul_ID, .keep_all = TRUE)
```





# Study Fleet and NEFSC Pairs
## 5m and 72hr Paired Tows
### Cod
#### Exploratory Vizualization 
```{r}
#Looking at the cpua for the paired tows for each species 
paired_tows_5mi_72hr_re %>%
  filter(SVSPP == 73 & SPECIES_ITIS == 164712) %>% 
  distinct(EFFORT_ID, .keep_all = TRUE) %>%
  filter(sf_species_cpua < 10000) %>%
  filter(nefsc_species_cpua < 10000) %>%
  ggplot() + 
  geom_bar(aes(x = EFFORT_ID, y = sf_species_cpua), stat = "identity", fill = "red") + 
  geom_bar(aes(x = EFFORT_ID, y = -nefsc_species_cpua), stat = "identity", fill = "blue") + 
  labs(x = "Tow Number", y = "CPUA (lbs/km2)", 
       title = "Comparison of CPUA for Atlantic Cod", 
       subtitle = "Pairs within 5 mi and 72 hrs")

#Looking at the relative efficiency of the study fleet tows compared to the nefsc survey 
paired_tows_5mi_72hr_re %>%
  filter(SVSPP == 73 & SPECIES_ITIS == 164712) %>%
  distinct(EFFORT_ID, .keep_all = TRUE) %>%
  filter(sf_species_cpua < 10000) %>%
  filter(nefsc_species_cpua < 10000) %>%
  ggplot() +
  geom_histogram(aes(x = relative_efficiency), bins = 50, color = "black", fill = "white") + 
  geom_vline(aes(xintercept = 0.5),
            color = "black", linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = mean(relative_efficiency)),
            color = "red", linetype = "dashed", size = 1) + 
  labs(x = "Relative Efficiency", y = "Count", 
       title = "Histogram of Relative Efficiency for Atlantic Cod", 
       subtitle = "Pairs within 5 mi and 72 hrs")
```

#### T-Test
```{r}
#looking at the boxplot and histograms of the catch per unit area 
box.plot(cod_paired_tows_5mi_72hr_re)
hist_kh(cod_paired_tows_5mi_72hr_re)

#log-transforming the data
cod_paired_tows_5mi_72hr_re <- cod_paired_tows_5mi_72hr_re %>%
  mutate(log_sf_species_cpua = log(sf_species_cpua), 
         log_nefsc_species_cpua = log(nefsc_species_cpua))

#looking at the boxplot and histograms of the log(catch per unit area)
box.log.plot(cod_paired_tows_5mi_72hr_re)
hist_log_kh(cod_paired_tows_5mi_72hr_re)

#assumptions of normality and equal variance are met with the log transformed data, 
#so the t-test was conducted with the log-transformed data


#running t-test
cod_ttest <- t.test(cod_paired_tows_5mi_72hr_re$log_sf_species_cpua, cod_paired_tows_5mi_72hr_re$log_nefsc_species_cpua, paired = TRUE)
cod_ttest


#back transform the estimate
exp(cod_ttest$estimate)
exp(cod_ttest$conf.int)

```

#### Model Runs
```{r}
#scatterplot matrix of all the variables of interest 
ggpairs(cod_paired_tows_5mi_72hr_re, columns = c("VESSEL_NAME", "season", "mgmt_area", "AVGDEPTH", "BOTTEMP"), 
        cardinality_threshold = 17)

#running all the models w every combination of variables and comparing them based on AIC and deviance explained 
codm <- fit.models(cod_paired_tows_5mi_72hr_re)
compare.models(codm)


#dropping the NAs in the bottom-temp column 
cod_paired_tows_5mi_72hr_re1 <- cod_paired_tows_5mi_72hr_re %>%
  drop_na(BOTTEMP)


#storing the optimal model and plotting the diagnostics
codm_optimal <- glm(relative_efficiency ~ VESSEL_NAME + season + BOTTEMP, family = binomial, data = cod_paired_tows_5mi_72hr_re1)
summary(codm_optimal)

par(mfrow=c(2,2))
plot(codm_optimal)




#back transform the estimates and CIs 
exp(cbind(OR = coef(codm_optimal), confint(codm_optimal)))

#look at the component residual plots 
crPlots(codm_optimal)

# looking at the effect_plot
effect_plot(codm_optimal, pred = VESSEL_NAME)




# Extract the fitted and predicted values 
cod_paired_tows_5mi_72hr_re1$predict <- predict(codm_optimal)
cod_paired_tows_5mi_72hr_re1$fitted <- fitted(codm_optimal)

#predicted vs observed values 
cod_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = fitted, 
             y = relative_efficiency)) + 
  geom_point() + 
  geom_abline(intercept=0, slope=1) +
  xlim(0,1) + 
  ylim(0,1) + 
  labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')

#predicted vs trawl type w point shape as explanatory variables 
#mgmt_area
cod_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = VESSEL_NAME, 
             y = fitted, color = mgmt_area)) + 
  geom_jitter() + 
  coord_flip()

#season
cod_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = VESSEL_NAME, 
             y = fitted, color = season)) + 
  geom_jitter() + 
  coord_flip()

#depth 
cod_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = VESSEL_NAME, 
             y = fitted, color = AVGDEPTH)) + 
  geom_jitter() + 
  coord_flip()






#getting final RE estimate 
cod_paired_tows_5mi_72hr_re1 %>%
  group_by(VESSEL_NAME) %>%
  summarise(mean(fitted), n = n())


cod_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = VESSEL_NAME, 
             y = fitted)) + 
  geom_jitter() + 
  stat_summary(
    geom = "point",
    fun = "mean",
    col = "black",
    size = 3,
    shape = 24,
    fill = "red"
  ) + 
  coord_flip()




# REGRESSION ANALYSIS USING BACKWARD ELIMINATION
#fit a full model 
cod_glm1 <- glm(relative_efficiency ~ VESSEL_NAME + season + mgmt_area + AVGDEPTH + BOTTEMP, 
                family = binomial, data = cod_paired_tows_5mi_72hr_re)
summary(cod_glm1)





```


### YellowTail Flounder
#### Exploratory Vizualization 
```{r}
#Looking at the cpua for the paired tows for each species 
paired_tows_5mi_72hr_re %>%
  filter(SVSPP == 105 & SPECIES_ITIS == 172909) %>% 
  distinct(EFFORT_ID, .keep_all = TRUE) %>%
  ggplot() + 
  geom_bar(aes(x = EFFORT_ID, y = sf_species_cpua), stat = "identity", fill = "blue") + 
  geom_bar(aes(x = EFFORT_ID, y = -nefsc_species_cpua), stat = "identity", fill = "red") + 
  labs(x = "Tow Number", y = "CPUA (lbs/km2)", 
       title = "Comparison of CPUA for Yellowtail Flounder", 
       subtitle = "Pairs within 5 mi and 72 hrs")

#Looking at the relative efficiency of the study fleet tows compared to the nefsc survey 
paired_tows_5mi_72hr_re %>%
  filter(SVSPP == 105 & SPECIES_ITIS == 172909) %>%
  distinct(EFFORT_ID, .keep_all = TRUE) %>%
  ggplot() +
  geom_histogram(aes(x = relative_efficiency), bins = 50, color = "black", fill = "white") + 
  geom_vline(aes(xintercept = 0.5),
            color = "black", linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = mean(relative_efficiency)),
            color = "red", linetype = "dashed", size = 1) + 
  labs(x = "Relative Efficiency", y = "Count", 
       title = "Histogram of Relative Efficiency for Yellowtail Flounder", 
       subtitle = "Pairs within 5 mi and 72 hrs")
```

#### T-Test
```{r}
#looking at the boxplot and histograms of the catch per unit area 
box.plot(yellowtailfl_paired_tows_5mi_72hr_re)
hist_kh(yellowtailfl_paired_tows_5mi_72hr_re)

#log-transforming the data
yellowtailfl_paired_tows_5mi_72hr_re <- yellowtailfl_paired_tows_5mi_72hr_re %>%
  mutate(log_sf_species_cpua = log(sf_species_cpua), 
         log_nefsc_species_cpua = log(nefsc_species_cpua))

#looking at the boxplot and histograms of the log(catch per unit area)
box.log.plot(yellowtailfl_paired_tows_5mi_72hr_re)
hist_log_kh(yellowtailfl_paired_tows_5mi_72hr_re)

#assumptions of normality and equal variance are not met with the log transformed data, 
#so the t-test was conducted with the log-transformed data


yellowtail_ttest <- t.test(yellowtailfl_paired_tows_5mi_72hr_re$log_sf_species_cpua, yellowtailfl_paired_tows_5mi_72hr_re$log_nefsc_species_cpua, 
                           paired = TRUE)
yellowtail_ttest


#back transform the estimate
exp(yellowtail_ttest$estimate)
exp(yellowtail_ttest$conf.int)

```

#### Model Runs 
```{r}
#scatterplot matrix of all the variables of interest 
ggpairs(yellowtailfl_paired_tows_5mi_72hr_re, columns = c("VESSEL_NAME", "season", "mgmt_area", "AVGDEPTH", "BOTTEMP"), 
        cardinality_threshold = 17)


#running all the models w every combination of variables and comparing them based on AIC and deviance explained 
yellowtailm <- fit.models(yellowtailfl_paired_tows_5mi_72hr_re)
compare.models(yellowtailm)


#dropping the NAs in the bottom-temp column 
yellowtailfl_paired_tows_5mi_72hr_re1 <- yellowtailfl_paired_tows_5mi_72hr_re %>%
  drop_na(BOTTEMP)



#storing the optimal model and plotting the diagnostics
yellowtailm_optimal <- glm(relative_efficiency ~ VESSEL_NAME + AVGDEPTH + BOTTEMP, family = binomial, data = yellowtailfl_paired_tows_5mi_72hr_re1)
summary(yellowtailm_optimal)

par(mfrow=c(2,2))
plot(codm_optimal)





# Extract the fitted and predicted values 
yellowtailfl_paired_tows_5mi_72hr_re1$predict <- predict(yellowtailm_optimal)
yellowtailfl_paired_tows_5mi_72hr_re1$fitted <- fitted(yellowtailm_optimal)

#predicted vs observed values 
yellowtailfl_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = fitted, 
             y = relative_efficiency)) + 
  geom_point() + 
  geom_abline(intercept=0, slope=1) +
  xlim(0,1) + 
  ylim(0,1) + 
  labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')

#predicted vs trawl type w point shape as explanatory variables 
#mgmt_area
yellowtailfl_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = VESSEL_NAME, 
             y = fitted, color = mgmt_area)) + 
  geom_jitter() + 
  coord_flip()

#season
yellowtailfl_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = VESSEL_NAME, 
             y = fitted, color = season)) + 
  geom_jitter() +  
  coord_flip()

#depth 
yellowtailfl_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = VESSEL_NAME, 
             y = fitted, color = AVGDEPTH)) + 
  geom_jitter() + 
  coord_flip()




#getting final RE estimate 
yellowtailfl_paired_tows_5mi_72hr_re1 %>%
  group_by(VESSEL_NAME) %>%
  summarise(mean(fitted), n = n())


yellowtailfl_paired_tows_5mi_72hr_re1 %>%
  ggplot(aes(x = VESSEL_NAME, 
             y = fitted)) + 
  geom_jitter() + 
  stat_summary(
    geom = "point",
    fun = "mean",
    col = "black",
    size = 3,
    shape = 24,
    fill = "red"
  ) + 
  coord_flip()


```

***

# Observer and NEFSC Pairs
## 5m and 72hr Paired Tows
### Cod
#### Exploratory Vizualization 
```{r}
paired_tows_5mi_72hr_ob_re %>%
  filter(SVSPP == 73 & NESPP4 == 818) %>% 
  distinct(trip_haul_ID, .keep_all = TRUE) %>%
  ggplot() + 
  geom_bar(aes(x = trip_haul_ID, y = ob_species_cpua), stat = "identity", fill = "blue") + 
  geom_bar(aes(x = trip_haul_ID, y = -nefsc_species_cpua), stat = "identity", fill = "red") + 
  labs(x = "Tow Number", y = "CPUA (lbs/km2)", 
       title = "Comparison of CPUA for Atlantic Cod", 
       subtitle = "Pairs within 5 mi and 72 hrs")

#Looking at the relative efficiency of the study fleet tows compared to the nefsc survey 
paired_tows_5mi_72hr_ob_re %>%
  filter(SVSPP == 73 & NESPP4 == 818) %>%
  distinct(trip_haul_ID, .keep_all = TRUE) %>%
  ggplot() +
  geom_histogram(aes(x = relative_efficiency), bins = 100, color = "black", fill = "white") + 
  geom_vline(aes(xintercept = 0.5),
            color = "black", linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = mean(relative_efficiency)),
            color = "red", linetype = "dashed", size = 1) + 
  labs(x = "Relative Efficiency", y = "Count", 
       title = "Histogram of Relative Efficiency for Atlantic Cod", 
       subtitle = "Pairs within 5 mi and 72 hrs")
```

#### T-Tests
```{r}
#looking at the boxplot and histograms of the catch per unit area 
box.plot_ob(cod_paired_tows_5mi_72hr_ob_re)
hist_kh_ob(cod_paired_tows_5mi_72hr_ob_re)

#log-transforming the data
cod_paired_tows_5mi_72hr_ob_re <- cod_paired_tows_5mi_72hr_ob_re %>%
  mutate(log_ob_species_cpua = log(ob_species_cpua), 
         log_nefsc_species_cpua = log(nefsc_species_cpua))

#looking at the boxplot and histograms of the log(catch per unit area)
box.log.plot_ob(cod_paired_tows_5mi_72hr_ob_re)
hist_log_kh_ob(cod_paired_tows_5mi_72hr_ob_re)

#assumptions of normality and equal variance are met with the log transformed data, 
#so the t-test was conducted with the log-transformed data

cod_ttest_ob <- t.test(cod_paired_tows_5mi_72hr_ob_re$log_ob_species_cpua, cod_paired_tows_5mi_72hr_ob_re$log_nefsc_species_cpua,
                              paired = TRUE)
cod_ttest_ob


#back transform the estimate
exp(cod_ttest_ob$estimate)
exp(cod_ttest_ob$conf.int)

```

#### Model Runs 
```{r}
#scatterplot matrix of all the variables of interest 
ggpairs(cod_paired_tows_5mi_72hr_ob_re, columns = c("trawl_type", "season", "mgmt_area", "AVGDEPTH", "BOTTEMP"))

#running all the models w every combination of variables and comparing them based on AIC and deviance explained 
codm <- fit.models_ob(cod_paired_tows_5mi_72hr_ob_re)
compare.models_ob(codm)


#storing the optimal model and plotting the diagnostics
codm_ob_optimal <- glm(relative_efficiency ~ trawl_type + AVGDEPTH, family = binomial, data = cod_paired_tows_5mi_72hr_ob_re)
summary(codm_ob_optimal)

par(mfrow=c(2,2))
plot(codm_ob_optimal)


#back transform the estimates and CIs 
exp(cbind(OR = coef(codm_ob_optimal), confint(codm_ob_optimal)))

#look at the component residual plots 
crPlots(codm_ob_optimal)

# looking at the effect_plot
effect_plot(codm_ob_optimal, pred = trawl_type, interval = TRUE, partial.residuals = TRUE, 
            jitter = 0.2)






# Extract the fitted and predicted values 
cod_paired_tows_5mi_72hr_ob_re$predict <- predict(codm_ob_optimal)
cod_paired_tows_5mi_72hr_ob_re$fitted <- fitted(codm_ob_optimal)

#predicted vs observed values 
cod_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = fitted, 
             y = relative_efficiency)) + 
  geom_point() + 
  geom_abline(intercept=0, slope=1) +
  xlim(0,1) + 
  ylim(0,1) + 
  labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')

#predicted vs trawl type w point shape as explanatory variables 
#mgmt_area
cod_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = trawl_type, 
             y = fitted, color = mgmt_area)) + 
  geom_jitter()

#season
cod_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = trawl_type, 
             y = fitted, color = season)) + 
  geom_jitter()

#depth 
cod_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = trawl_type, 
             y = fitted, color = AVGDEPTH)) + 
  geom_jitter()




#getting final RE estimate 
cod_paired_tows_5mi_72hr_ob_re %>%
  group_by(trawl_type) %>%
  summarise(mean(fitted), n = n())


cod_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = trawl_type, 
             y = fitted)) + 
  geom_jitter() + 
  stat_summary(
    geom = "point",
    fun = "mean",
    col = "black",
    size = 3,
    shape = 24,
    fill = "red"
  )


```




### YellowTail Flounder
#### Exploratory Vizualization 
```{r}
#Looking at the cpua for the paired tows for each species 
paired_tows_5mi_72hr_ob_re %>%
  filter(SVSPP == 105 & NESPP4 == 1230) %>% 
  distinct(trip_haul_ID, .keep_all = TRUE) %>%
  ggplot() + 
  geom_bar(aes(x = trip_haul_ID, y = ob_species_cpua), stat = "identity", fill = "blue") + 
  geom_bar(aes(x = trip_haul_ID, y = -nefsc_species_cpua), stat = "identity", fill = "red") + 
  labs(x = "Tow Number", y = "CPUA (lbs/km2)", 
       title = "Comparison of CPUA for Yellowtail Flounder", 
       subtitle = "Pairs within 5 mi and 72 hrs")

#Looking at the relative efficiency of the study fleet tows compared to the nefsc survey 
paired_tows_5mi_72hr_ob_re %>%
  filter(SVSPP == 105 & NESPP4 == 1230) %>%
  distinct(trip_haul_ID, .keep_all = TRUE) %>%
  ggplot() +
  geom_histogram(aes(x = relative_efficiency), bins = 100, color = "black", fill = "white") + 
  geom_vline(aes(xintercept = 0.5),
            color = "black", linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = mean(relative_efficiency)),
            color = "red", linetype = "dashed", size = 1) + 
  labs(x = "Relative Efficiency", y = "Count", 
       title = "Histogram of Relative Efficiency for Yellowtail Flounder", 
       subtitle = "Pairs within 5 mi and 72 hrs")
```

#### T-Tests
```{r}
#looking at the boxplot and histograms of the catch per unit area 
box.plot_ob(yellowtailfl_paired_tows_5mi_72hr_ob_re)
hist_kh_ob(yellowtailfl_paired_tows_5mi_72hr_ob_re)

#log-transforming the data
yellowtailfl_paired_tows_5mi_72hr_ob_re <- yellowtailfl_paired_tows_5mi_72hr_ob_re %>%
  mutate(log_ob_species_cpua = log(ob_species_cpua), 
         log_nefsc_species_cpua = log(nefsc_species_cpua))

#looking at the boxplot and histograms of the log(catch per unit area)
box.log.plot_ob(yellowtailfl_paired_tows_5mi_72hr_ob_re)
hist_log_kh_ob(yellowtailfl_paired_tows_5mi_72hr_ob_re)

#assumptions of normality and equal variance are met with the log transformed data, 
#so the t-test was conducted with the log-transformed data

yellowtail_ttest_ob <- t.test(yellowtailfl_paired_tows_5mi_72hr_ob_re$log_ob_species_cpua, yellowtailfl_paired_tows_5mi_72hr_ob_re$log_nefsc_species_cpua,
                              paired = TRUE)
yellowtail_ttest_ob


#back transform the estimate
exp(yellowtail_ttest_ob$estimate)
exp(yellowtail_ttest_ob$conf.int)
```

#### Model Runs 
```{r}
#scatterplot matrix of all the variables of interest 
ggpairs(yellowtailfl_paired_tows_5mi_72hr_ob_re, columns = c("trawl_type", "season", "mgmt_area", "AVGDEPTH", "BOTTEMP"))

#running all the models w every combination of variables and comparing them based on AIC and deviance explained 
yellowtailm_ob <- fit.models_ob(yellowtailfl_paired_tows_5mi_72hr_ob_re)
compare.models_ob(yellowtailm_ob)


#storing the optimal model and plotting the diagnostics
yellowtailm_ob_optimal <- glm(relative_efficiency ~ trawl_type + season, family = binomial, data = yellowtailfl_paired_tows_5mi_72hr_ob_re)
summary(yellowtailm_ob_optimal)

par(mfrow=c(2,2))
plot(yellowtailm_ob_optimal)


#back transform the estimates and CIs 
exp(cbind(OR = coef(yellowtailm_ob_optimal), confint(yellowtailm_ob_optimal)))

#look at the component residual plots 
crPlots(yellowtailm_ob_optimal)

# looking at the effect_plot
effect_plot(yellowtailm_ob_optimal, pred = trawl_type, interval = TRUE, partial.residuals = TRUE, 
            jitter = 0.2)




# Extract the fitted and predicted values 
yellowtailfl_paired_tows_5mi_72hr_ob_re$predict <- predict(yellowtailm_ob_optimal)
yellowtailfl_paired_tows_5mi_72hr_ob_re$fitted <- fitted(yellowtailm_ob_optimal)


#predicted vs observed values 
yellowtailfl_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = fitted, 
             y = relative_efficiency)) + 
  geom_point() + 
  geom_abline(intercept=0, slope=1) +
  xlim(0,1) + 
  labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')
  


#predicted vs trawl type w point shape as explanatory variables 
#mgmt_area
yellowtailfl_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = trawl_type, 
             y = fitted, color = mgmt_area)) + 
  geom_jitter()

#season
yellowtailfl_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = trawl_type, 
             y = fitted, color = season)) + 
  geom_jitter()

#depth 
yellowtailfl_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = trawl_type, 
             y = fitted, color = AVGDEPTH)) + 
  geom_jitter()



#getting final RE estimate 
yellowtailfl_paired_tows_5mi_72hr_ob_re %>%
  group_by(trawl_type) %>%
  summarise(mean(fitted), n = n())


yellowtailfl_paired_tows_5mi_72hr_ob_re %>%
  ggplot(aes(x = trawl_type, 
             y = fitted)) + 
  geom_jitter() + 
  stat_summary(
    geom = "point",
    fun = "mean",
    col = "black",
    size = 3,
    shape = 24,
    fill = "red"
  )

```





