---
title: "Hankowsky_HW8"
author: "Keith Hankowsky"
date: '2022-10-25'
output: html_document
---


```{r setup, include=FALSE}
library(Sleuth3)
library(tidyverse)
library(asht)
library(mosaic)
library(GGally)
library(gridExtra)
library(olsrr)
library(car)

```

Hankowsky Homework Solutions 


# Basic Problems 
## 13.16
Toxic Effects of Copper and Zinc. Reconsider the data in Display 10.20 (file: ex1014). (a) Is
the experiment a randomized block design or a completely randomized design with factorial treatment
structure? (b) Analyze the data using two-way analysis of variance. (c) How does the analysis
of variance compare to the regression analysis (i.e., what are the issues in deciding which analysis is
more appropriate)?
```{r}
#Read-In the data
minnow <- Sleuth3::ex1014

#take a look at the dataset 
head(minnow)

#turn copper and zinc into factors 
minnow <- minnow %>%
  mutate(Copper = as.factor(Copper), 
         Zinc = as.factor(Zinc))


#look at the boxplots 
p1 <- minnow %>%
  ggplot() + 
  geom_boxplot(aes(x = as.factor(Copper), y = Protein)) + 
  labs(x = "Copper") + 
  theme_classic()

p2 <- minnow %>%
  ggplot() + 
  geom_boxplot(aes(x = as.factor(Zinc), y = Protein)) + 
  labs(x = "Zinc", y = "") + 
  theme_classic()

grid.arrange(p1, p2, nrow = 1)


#look at the interaction plot
with(minnow, interaction.plot(Copper, Zinc, Protein))


# REGRESSION ANALYSIS
#running the full model 
minnow_lm <- lm(Protein ~ Copper + Zinc, data = minnow) #when I included the interaction term it wasn't really running the subsequent code well 
anova(minnow_lm)

par(mfrow=c(2,2))
plot(minnow_lm)

summary(minnow_lm)


# ANOVA ANALYSIS 
minnow_aov <- aov(Protein ~ Copper + Zinc, data = minnow)

par(mfrow=c(2,2))
plot(minnow_aov)

summary(minnow_aov)


```

The experiment is a randomized block design, because the beakers containing minnows were randomly allocated to one of the treatments, each of which constitutes a single trial. 


***


## 13.18
El Ni ˜no and Hurricanes. Reconsider the El Ni˜no and Hurricane data set from Exercise 10.28.
(a) Regress the log of the storm index on West African wetness (treated as a categorical factor with
2 levels) and El Ni˜no temperature (treated as a categorical factor with 3 levels); retain the sum of
squared residuals and the residual degrees of freedom. (b) Regress the log of the storm index onWest
African wetness (treated as categorical with 2 levels), El Ni˜no temperature (treated as numerical),
and the square of El Ni˜no temperature. Retain the sum of squared residuals and the residual degrees
of freedom. (c) Explain why the answers to (a) and (b) are the same. (d) Explain why a test that
the coefficient of the temperature-squared term is zero can be used to help decide whether to treat
temperature as numerical or categorical.
```{r}
#Read-In the data
hurricane <- Sleuth3::ex1028

#take a look at the dataset 
head(hurricane)

#transform the data as described in the question 
hurricane <- hurricane %>%
  mutate(log_stormindex = log(StormIndex), 
         Temperature_a = as.factor(Temperature), 
         WestAfrica_cat = as.factor(WestAfrica))



# REGRESSION ANALYSIS - PART A
#run the regression 
hurricane_a_lm1 <- lm(log_stormindex ~ WestAfrica_cat + Temperature_a + WestAfrica_cat:Temperature_a, data = hurricane)

par(mfrow=c(2,2))
plot(hurricane_a_lm1)

#refit without interaction and test for interaction effect
hurricane_a_lm2 <- update(hurricane_a_lm1, ~ . - WestAfrica_cat:Temperature_a)
anova(hurricane_a_lm2, hurricane_a_lm1) # no evidence for an interaction effect, p-value = 0.56


#run optimal model, question asks for both West Africa wetness and El Nino temp to be in the model 
hurricane_a_lm <- lm(log_stormindex ~ WestAfrica_cat + Temperature_a, data = hurricane)

summary(hurricane_a_lm)
anova(hurricane_a_lm)

par(mfrow=c(2,2))
plot(hurricane_a_lm)



# REGRESSION ANALYSIS - PART B 
hurricane_b_lm1 <- lm(log_stormindex ~ WestAfrica_cat + Temperature + WestAfrica_cat:Temperature, data = hurricane)

par(mfrow=c(2,2))
plot(hurricane_b_lm1)

#refit without interaction and test for interaction effect
hurricane_b_lm2 <- update(hurricane_b_lm1, ~ . - WestAfrica_cat:Temperature)
anova(hurricane_b_lm2, hurricane_b_lm1) # no evidence for an interaction effect, p-value = 0.60


#run optimal model, question asks for both West Africa wetness and El Nino temp to be in the model 
hurricane_b_lm <- lm(log_stormindex ~ WestAfrica_cat + Temperature, data = hurricane)

summary(hurricane_b_lm)
anova(hurricane_b_lm)

par(mfrow=c(2,2))
plot(hurricane_b_lm)



# PART C - Compare the SS residuals and residual df
a1 <- anova(hurricane_a_lm)
a2 <- anova(hurricane_b_lm)

#extracting the SS residuals 
a1["Residuals", "Sum Sq"]
a2["Residuals", "Sum Sq"]

#extracting the residual df
a1["Residuals", "Df"]
a2["Residuals", "Df"]




```





***

# Supplemental Problems 
## 13.17
Dinosaur Extinctions—An Observational Study. About 65 million years ago, the dinosaurs
suffered a mass extinction virtually overnight (in geologic time).What happened in this period, the
Cretaceous–Tertiary (KT) boundary, that could have produced such calamity? Among many clues,
one that all scientists regard as crucial is a layer of iridium-rich dust that was deposited over much of
the earth at that time. Data from one of the first discoveries of this layer are shown inDisplay 13.22.
The diagram traces iridium concentrations in soil samples taken from extensive shale and limestone
deposits at Gubbio, Italy. Iridium (parts per trillium) is graphed against the depth at which samples
were taken, with depth giving a crude measure of historic time.

Iridium is a trace element present in most soils. But concentrations as high as those at the peak
near 347 meters, at the KT boundary, can only occur in association with highly unusual events, such
as volcanic eruptions or meteor impacts. So the theory is that such an event caused a massive dust
cloud that blanketed the earth for years, killing off animals and their food sources. Butwas the cause a
meteor (coming down on the Yucatan peninsula in central America) or volcanic eruptions (centered
in southern China)? Two articles debating the issue appeared in the October 1990 issue of Scientific
American—W. Alvarez and F. Asaro, “What Caused the Mass Extinction? An Extraterrestrial
Impact,” Scientific American 263(4): 76–84, and E. Courtillot, “What Caused the Mass Extinction?
A Volcanic Eruption,” Scientific American 263(4): 85–93. A crucial issue in the debate is the shape
of the iridium trace because the timing and extent of the source give clues to its nature.

The raw data are provided in Display 13.23. (a) Fit the two-way, saturated model to the untransformed
data and draw a plot of the residuals versus estimated means to see if a transformation
is suggested. (b) Fit the two-way model (after transformation if appropriate) and test for interaction,
using a multiple regression routine. (c) If appropriate with your statistical software, repeat part
(b) using an analysis of variance routine.
```{r}
#Read-In the data
dinos <- Sleuth3::ex1317

#take a look at the dataset 
head(dinos)

#transforming the data for later in the problem 
dinos <- dinos %>%
  mutate(log_iridium = log(Iridium))

#take a look at the boxplots 
p31 <- dinos %>%
  ggplot() + 
  geom_boxplot(aes(x = Strata, y = Iridium)) + 
  theme_classic()

p32 <- dinos %>%
  ggplot() + 
  geom_boxplot(aes(x = DepthCat, y = Iridium)) + 
  labs(x = "Depth Category", y = "") + 
  theme_classic()

grid.arrange(p31, p32, nrow = 1)



# PART A 
dinos_a_lm <- lm(Iridium ~ Strata + DepthCat, data = dinos)

par(mfrow=c(2,2))
plot(dinos_a_lm)



# PART B 
dinos_b_lm1 <- lm(log_iridium ~ Strata + DepthCat + Strata:DepthCat, data = dinos)

par(mfrow=c(2,2))
plot(dinos_b_lm)


#refit without interaction and test for interaction effect
dinos_b_lm2 <- update(dinos_b_lm1, ~ . - Strata:DepthCat)
anova(hurricane_b_lm2, hurricane_b_lm1) # no evidence for an interaction effect, p-value = 0.60


#run optimal model
dinos_b_lm <- lm(log_iridium ~ Strata + DepthCat, data = dinos)

summary(dinos_b_lm)
anova(dinos_b_lm)

par(mfrow=c(2,2))
plot(dinos_b_lm)





# PART C 

```




***


##13.21
Pygmalion. The term Pygmalion effect (used in the Section 13.1.2 data problem) originated
with the 1960s’ work of psychologist Robert Rosenthal and elementary school principal Lenore
Jacobson, who conducted a randomized block experiment on the children in Jacobson’s elementary
school. In each classroom (i.e., for each block of students), they assigned roughly 20% of the
students to a “likely to excel” treatment group and the remainder to a control group. After all students
took a standardized test at the begining of the school year, Rosenthal and Jacobson identified
to teachers those students that had been indicated by the intelligence test as very likely to excel in
the coming year. This was false information, though. The “likely to excel” students were simply
those who had been assigned by a chance mechanism to the “likely to excel” treatment group. The
researchers deceived the teachers with this false information to create artificial expectations and to
explore the effect of those expectations on student test score gains. Display 13.26 is a partial listing
of a data set simulated to match the summary statistics and conclusions from Table A-7 of Rosenthal
and Jacobson’s report (R. Rosenthal, and L. Jacobson, 1968, Pygmalion in the Classroom: Teacher
Expectation and Pupil’s Intellectual Development, Holt, Rinehart, andWinston, Inc.). Analyze the
data to summarize the evidence that teachers’ expectations affect student test score gains. Write a
statistical report of your conclusions.
```{r}
#Read-In the data
pygmalion <- Sleuth3::ex1321

#take a look at the dataset 
head(pygmalion)


#take a look at the boxplots 
pygmalion %>%
  ggplot() + 
  geom_boxplot(aes(x = Treatment, y = Gain)) + 
  labs(title = "Boxplot of Student Test Score Gains vs Treatment")
  theme_classic()

pygmalion %>%
  ggplot() + 
  geom_boxplot(aes(x = Class, y = Gain, color = Treatment)) + 
  labs(title = "Boxplot of Student Test Score Gains vs Classroom") + 
  theme_classic()

#take a look at interaction plot 
with(pygmalion, interaction.plot(Class, Treatment, Gain))


# REGRESSION ANALYSIS
pygmalion_lm1 <- lm(Gain ~ Treatment + Class + Treatment:Class, data = pygmalion)

par(mfrow=c(2,2))
plot(pygmalion_lm1)

#refit without interaction and test for interaction effect
pygmalion_lm2 <- update(pygmalion_lm1, ~ . - Treatment:Class)
anova(pygmalion_lm2, pygmalion_lm1) # no evidence for an interaction effect, p-value = 0.11

#refit without class and test for class effect 
pygmalion_lm3 <- update(pygmalion_lm2, ~ . - Class)
anova(pygmalion_lm3, pygmalion_lm2) # strong evidence for a class effect, p-value < 0.0001


#run optimal model
pygmalion_lm <- lm(Gain ~ Treatment + Class, data = pygmalion)

summary(pygmalion_lm)
anova(pygmalion_lm)

par(mfrow=c(2,2))
plot(pygmalion_lm)

confint(pygmalion_lm)


#randomization distribution 
obs = summary(lm(Gain ~ Treatment + Class, data = pygmalion))$coefficients["Treatmentpygmalion", "t value"]

nulldist = do(15000) * summary(lm(Gain ~ shuffle(Treatment) + shuffle(Class), data = pygmalion))$coefficients["shuffle(Treatment)pygmalion", "t value"]

histogram(~result, groups = result >= obs, v = obs, data = nulldist)
tally(~result >= obs, format = "proportion", data = nulldist)

```

The Pygmalion treatments adds an estimated 3.61 points to a students test score (95% confidence interval: 0.147 to 12.90 points). The study provides moderate evidence that the effect is real (one-sided p-value = 0.021). 

